# RSiena analysis Computing science staff RU

<br>

## Preparations
```{r}
# Starting with a clean workspace 
rm(list=ls())
```

```{r}
# Requiring packages 
library(data.table) 
library(tidyverse) 
require(stringi)
require(RSiena)
```

```{r}
# Loading dataobjects
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\comp_net_array.RData")
load("C:\\Users\\Daphn\\Documenten\\pubnets\\Data\\names_staff_cit_v20221006.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\comp_df.RData")
```

```{r}
# Making the network undirected
net1 <- net_comp_array[,,1]
net1 <- net1 + t(net1)
net1[net1==2] <- 1

net2 <- net_comp_array[,,2]
net2 <- net2 + t(net2)
net2[net2==2] <- 1

net3 <- net_comp_array[,,3]
net3 <- net3 + t(net3)
net3[net3==2] <- 1


net_comp_array <- array(data = c(net1, net2, net3), dim=c(dim(net1),3))
```

```{r}
# Selecting the dependent variable
net <- sienaDependent(net_comp_array)
?sienaDependent
```

<br>

## Define data
```{r}
# Gender
gender <- as.numeric(comp_df$gender=="female")
gender <- coCovar(gender)
```

```{r}
# Ethnicity
dutch <- as.numeric(comp_df$dutch==1)
dutch <- coCovar(dutch)
```

```{r}
# Number of collaborators
#ncollabs <- comp_df$ncollabs
#ncollabs <- coCovar(ncollabs)

#class(ncollabs)
```

```{r}
# Creating a dataset with information on the google scholar id's, publications and citations of each computing science staff member 

names_staff_cit %>%
  filter(names_staff_cit$gs_id %in% comp_df$gs_id) -> comp_staff_cit

# Removing duplicates
sel <- (duplicated.data.frame(comp_staff_cit))

comp_staff_cit %>% 
  filter(!sel) -> comp_staff_cit
```

```{r}
# Publications per wave for each staff computing science staff member
pubsw1 <- pubsw2 <- pubsw3 <- NA

for (i in 1:length(comp_df$gs_id)) {
  pubsw1[i] <- nrow(comp_staff_cit[(comp_staff_cit$gs_id == comp_df$gs_id[i]) & comp_staff_cit$year>=2016 & comp_staff_cit$year<=2017,])
  pubsw2[i] <- nrow(comp_staff_cit[(comp_staff_cit$gs_id == comp_df$gs_id[i]) & comp_staff_cit$year>=2018 & comp_staff_cit$year<=2019,])
  pubsw3[i] <- nrow(comp_staff_cit[(comp_staff_cit$gs_id == comp_df$gs_id[i]) & comp_staff_cit$year>=2020 & comp_staff_cit$year<=2022,])
}

pub_df <- as.matrix(data.frame(pubsw1, pubsw2, pubsw3))

pubs <- varCovar(pub_df)
```

```{r, warning=FALSE, message=FALSE}
# Year of first publication
comp_staff_cit %>% group_by(gs_id) %>%
  mutate(pub_first = min(year)) %>%
  select(c("gs_id", "pub_first")) %>%
  distinct(gs_id, pub_first, .keep_all = TRUE) -> firstpub_df

comp_df <- left_join(comp_df, firstpub_df)

#if no publication yet, set pub_first op 2023
comp_df %>% mutate(pub_first = replace_na(pub_first, 2023)) -> comp_df

pub_first <-  coCovar(comp_df$pub_first)

mydata <- sienaDataCreate(net, gender, dutch, pubs, pub_first)
```

## Step 0: get initial description
```{r}
print01Report(mydata, modelname = "C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_init")
```

![](results/comp_init.txt){#id .class width=100% height=200px}

<br>

## Model 1

### Step 1: create effects structure
```{r}
myeff <- getEffects(mydata)
#effectsDocumentation(myeff)
```

<br>

### Step 2: specify model 1 (Only gender effects)
```{r}
# Gender effects
myeff <- includeEffects(myeff, egoPlusAltX, interaction1 = "gender")
myeff <- includeEffects(myeff, egoXaltX, interaction1 = "gender")
myeff
```

<br>

### Step 3: estimate
```{r}
myAlgorithm <- sienaAlgorithmCreate(projname = "comp_init", )
```

```{r, eval=FALSE}
(ans1_a <- siena07(myAlgorithm, nbrNodes=7, data = mydata, effects = myeff))
(ans1_b <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns = ans1_a))

save(ans1..., file="C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_model1")
```

```{r, echo=FALSE}
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_model1")
summary(ans1...)
```


<br>

## Model 2

### Step 1: create effects structure
```{r}
myeff <- getEffects(mydata)
#effectsDocumentation(myeff)
```

<br>

### Step 2: specify model 2 (structural effects)
```{r}
# Structural effects
myeff <- includeEffects(myeff, density)
#myeff <- includeEffects(myeff, isolateNet) #lijkt tot convergentie problemen te leiden. wellicht fixen op een hoge waarde??? zie manual als je ooit nog tijd hebt. 
#myeff <- includeEffects(myeff, transTrip) #o, niet goed gekeken, deze alleen voor directed netwerken
myeff <- includeEffects(myeff, transTies)
myeff <- includeEffects(myeff, degPlus)
```

```{r, echo=FALSE}
# Possible structural effects:
# density
# transTrip, transTies, gwesp, 
# inPop, outAct, inAct of outPop, 
# isolateNet
# degPlus of egoPlusAltX
```

<br>

### Step 3: estimate
```{r}
myAlgorithm <- sienaAlgorithmCreate(projname = "comp_init", )
```

```{r, eval=FALSE}
(ans2_a <- siena07(myAlgorithm, nbrNodes=7, data = mydata, effects = myeff))
(ans2_b <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns = ans2_a))

save(ans2..., file="C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_model2")
```

```{r, echo=FALSE}
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_model2")
summary(ans2...)
```

<br>

## Model 3

### Step 1: create effects structure
```{r}
myeff <- getEffects(mydata)
#effectsDocumentation(myeff)
```

<br>

### Step 2: specify model 3 (structural effects and gender effects)
```{r}
# Structural effects
myeff <- includeEffects(myeff, density)
myeff <- includeEffects(myeff, transTies)
myeff <- includeEffects(myeff, degPlus)
```

```{r}
# Gender effects
myeff <- includeEffects(myeff, egoPlusAltX, interaction1 = "gender")
myeff <- includeEffects(myeff, egoXaltX, interaction1 = "gender")
myeff
```

<br>

### Step 3: estimate
```{r}
myAlgorithm <- sienaAlgorithmCreate(projname = "comp_init", )
```

```{r, eval=FALSE}
(ans3_a <- siena07(myAlgorithm, nbrNodes=7, data = mydata, effects = myeff))
(ans3_b <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns = ans3_a))

save(ans3..., file="C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_model2")
```

```{r, echo=FALSE}
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_model3")
summary(ans3...)
```

<br>

## Model 4

### Step 1: create effects structure
```{r}
myeff <- getEffects(mydata)
#effectsDocumentation(myeff)
```

<br>

### Step 2: specify model 3 (structural effects, gender effects and covariate effects)
```{r}
# Structural effects
myeff <- includeEffects(myeff, density)
myeff <- includeEffects(myeff, transTies)
myeff <- includeEffects(myeff, degPlus)
```

```{r}
# Gender effects and covariate effects
myeff <- includeEffects(myeff, sameX, interaction1 = "dutch")
myeff <- includeEffects(myeff, egoX, interaction1 = "pub_first")
myeff <- includeEffects(myeff, egoX, interaction1 = "dutch")
myeff <- includeEffects(myeff, egoPlusAltX, interaction1 = "gender")
myeff <- includeEffects(myeff, egoXaltX, interaction1 = "gender")
myeff
```

```{r, echo=FALSE}
# Possible gender and covariate effects
#myeff <- includeEffects(myeff, sameX, interaction1 = "gender")
#myeff <- includeEffects(myeff, sameX, interaction1 = "dutch")
#myeff <- includeEffects(myeff, egoX, interaction1 = "pub_first")
#myeff <- includeEffects(myeff, egoX, interaction1 = "gender")
#myeff <- includeEffects(myeff, altX, interaction1 = "gender")
#myeff <- includeEffects(myeff, egoPlusAltX, interaction1 = "gender")
#myeff <- includeEffects(myeff, egoXaltX, interaction1 = "gender")
#myeff <- includeEffects(myeff, egoX, interaction1 = "dutch")
#myeff
```

<br>

### Step 3: estimate
```{r}
myAlgorithm <- sienaAlgorithmCreate(projname = "comp_init", )
```

```{r, eval=FALSE}
(ans4_a <- siena07(myAlgorithm, nbrNodes=7, data = mydata, effects = myeff))
(ans4_b <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns = ans4_a))

save(ans4..., file="C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_model4")
```

```{r, echo=FALSE}
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_model4")
summary(ans4...)
```
