# getting started

```{r}
#start with clean workspace 
rm(list=ls())
getwd()
```

# packages
```{r}
#install.packages("data.table") 
library(data.table) # mainly for faster data handling
library(tidyverse) # I assume you already installed this one!
# install.packages("httr") # we don't need this for now
# require(httr)
#install.packages("xml2")
require(xml2)
#install.packages("rvest")
require(rvest)
#install.packages("devtools")
require(devtools)
# Note we're doing something different here. We're installing a *latest* version directly from GitHub
# This is because the released version of this packages contains some errors!
#devtools::install_github("jkeirstead/scholar") 


require(scholar)

#define workdirectory, note the double *backslashes* if you're on windows
# setwd("/yourpathhere)"
```

# save webpage with staff names 

```{r}
# Let's first get the staff page read_html is a function that simply extracts html webpages and
# puts them in xml format
cs_staff <- read_html("https://www.cs.ru.nl/staff/index.html")

#head(soc_staff)
#class(soc_staff)
```

# from xml to nice df of names

## selecting the table with names

```{r}
cs_staff <- cs_staff %>%
    rvest::html_nodes("body") %>%
    xml2::xml_find_all("//a") %>%
    rvest::html_text()
```


```{r}
Encoding(cs_staff) <- "UTF-8"
cs_staff <- iconv(cs_staff, from="UTF-8", to="LATIN1")
require(stringi)
cs_staff <- stri_trans_general(cs_staff, id = "Latin-ASCII")

check <- NA

for (i in 1:length(cs_staff)) {
  check[i] <- str_length(cs_staff[i]) > 1
}

cs_staff2 <- cs_staff[check]
cs_staff <- cs_staff2
```


## names in df
```{r}
cs_df <- data.frame(cs_staff)  
```

Perhaps, you want to save `cs_df`. 

# Cleaning the dataset  

## cleaning names a bit  
```{r}
require(stringr)
# Last name seems to be everything after ) 
cs_df$last_name <-as.character(str_split(cs_df$cs_staff, pattern="\\)", n = 2, simplify = TRUE)[,2])

# first name is everything between brackets
cs_df$first_name <- as.character(str_extract_all(cs_df$cs_staff, "(?<=\\().+?(?=\\))", simplify = TRUE))
```

```{r}
cs_df$first_name <- tolower(cs_df$first_name)  # everything to lower!
cs_df$last_name <- tolower(cs_df$last_name)
```

Please note I updated below to loose double names
```{r}
# trimws looses all spacing before and after (if you specify 'both') a character string
cs_df$last_name <- trimws(cs_df$last_name, which = c("both"), whitespace = "[ \t\r\n]")
cs_df$first_name <- trimws(cs_df$first_name, which = c("both"), whitespace = "[ \t\r\n]")

cs_df$first_name <- as.character(str_split(cs_df$first_name, pattern=" ", n = 2, simplify = TRUE)[,1])
cs_df$first_name <- as.character(str_split(cs_df$first_name, pattern="-", n = 2, simplify = TRUE)[,1])

#cs_df$last_name <- as.character(str_split(cs_df$last_name, pattern=" ", n = 2, simplify = TRUE)[,1])
#cs_df$last_name <- as.character(str_split(cs_df$last_name, pattern="-", n = 2, simplify = TRUE)[,1])


#cs_df$soc_names <- trimws(cs_df$soc_names, which = c("both"), whitespace = "[ \t\r\n]")
cs_df$last_name
```


```{r}
save(cs_df, file="cs_df_s1b.RData")
```


## adding gender to dataset

### fgender

Hereyago. 

```{r}
fgender <- function(firstname_df, me, file=NULL) {

####################################
# Author: Bas Hofstra, Anne Maaike Mulders, Jochem Tolsma
# DAte:   13-10-2021, last edit: 22-09-2022
# Tasks:  - assign gender baed on name
#         - Adapted from Rense Corten code April 2021
####################################


#Input: 
#  - firstname_df: a data.frame with a column named firstname  and gender!
#  - me: a character vector introducing yourself: e.g. "J Tolsma, Radboud University"
#  - file: location and name of file to be saved. 
  
#------------------------------------------------------------------------------------
# Load required packages

if (!require("tidyverse", character.only = TRUE)) {
  install.packages("tidyverse", dependencies = TRUE)
  library(tidyverse, character.only = TRUE)
}

if (!require("rvest", character.only = TRUE)) {
  install.packages("rvest", dependencies = TRUE)
  library(rvest, character.only = TRUE)
}

if (!require("polite", character.only = TRUE)) {
  install.packages("polite", dependencies = TRUE)
  library(polite, character.only = TRUE)
}



# make links to scrape
firstname_df$name_url <- paste0("https://www.meertens.knaw.nl/nvb/naam/is/", firstname_df[, c("firstname")])



#------------------------------------------------------------------------------------
### 2: introduce to server ###

# Introduce myself to the server
session <- bow("https://www.meertens.knaw.nl/nvb/naam/is", user_agent = me , delay = 1)


#------------------------------------------------------------------------------------
### 3: make function to get table from ###
  fnames <- function(link){ 
    name_session <-nod(session, path = link)
    name_page <- scrape(name_session) 
    return(name_page)
  }
  
name_list <- list()
table_list <- list()


  for (i in 1:nrow(firstname_df)) {
    print(i)
    if (!(is.na(firstname_df$gender))) next
    name_list[[i]] <- fnames(firstname_df[i, c("name_url")])
    # extract name frequency table and gender info
    if (length(name_list[[i]] %>% html_table())>0) {
      
      table_list[[i]] <- name_list[[i]] %>% html_table()
      table_list[[i]][[1]][table_list[[i]][[1]]=="--"] <- "0"
      if (as.numeric(table_list[[i]][[1]]$X3[2]) > as.numeric(table_list[[i]][[1]]$X3[6])) {
        firstname_df$gender[i] <- "male" } else {
          firstname_df$gender[i] <- "female"
        }
    }
    if (!is.null(file)) (save(firstname_df, file=file))
    
    }
  return(firstname_df)
}
```




### gender to cs_df
```{r}
cs_df %>% mutate(firstname=first_name) -> cs_df
cs_df$gender <- NA
```


```{r}
cs_df$firstname
cs_df <- fgender(cs_df, me="Jochem Tolsma, RU/RUG", file="tempgender_cs_d2.RData")

```

```{r}
save(cs_df, file="cs_df_s2b.RData") #genderized last. 
```

## adding some other data manually

### affiliation
```{r}
# set affiliation to radboud, comes in handy for querying google scholar
cs_df$affiliation <- "radboud university"  
```

### other stuff

to do

# let harvest data from google scholar

## fixing a bug in the get_scholar_id function. 

```{r}
#require(scholar)

get_scholar_id_fix <- function (last_name = "", first_name = "", affiliation = NA)
{
  if (!any(nzchar(c(first_name, last_name))))
    stop("At least one of first and last name must be specified!")
  site <- getOption("scholar_site")
  url <- paste0(site, "/citations?view_op=search_authors&mauthors=",
                first_name, "+", last_name, "&hl=en&oi=ao")
  page <- get_scholar_resp(url)
  if (is.null(page))
    return(NA)
  aa <- httr::content(page, as = "text")
  # added by Bas Hofstra: bugfix for IDs that have a dash ("-")
  ids <- substring(aa, regexpr(";user=", aa))
  ids <- substr(ids, 1, 19) # error prone, but unsure how to solve otherwise
  # if (nchar(stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]")[[1]][1]) < 18) {
  #   ids <- stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]+[[:alnum:]]+[[:punct:]]")
  # } else {
  #   ids <- stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]")
  # }
  if (length(unlist(ids)) == 0) {
    message("No Scholar ID found.")
    return(NA)
  }
  ids <- ids %>% unlist %>% gsub(";user=|[[:punct:]]$", "",
                                 .) %>% unique
  if (length(ids) > 1) {
    profiles <- lapply(ids, scholar::get_profile)
    if (is.na(affiliation)) {
      x_profile <- profiles[[1]]
      warning("Selecting first out of ", length(profiles),
              " candidate matches.")
    }
    else {
      which_profile <- sapply(profiles, function(x) {
        stringr::str_count(string = x$affiliation, pattern = stringr::coll(affiliation,
                                                                           ignore_case = TRUE))
      })
      if (all(which_profile == 0)) {
        warning("No researcher found at the indicated affiliation.")
        return(NA)
      }
      else {
        x_profile <- profiles[[which(which_profile !=
                                       0)]]
      }
    }
  }
  else {
    x_profile <- scholar::get_profile(id = ids)
  }
  return(x_profile$id)
}
```

## scholars id. 

Don't forget to check manually if everything is okay. 

We make an important decisson here. we remove staff members without scholar ids. perhaps a bit strange for our RSiena analysis later. I mean, we can include isolates

YOU MAY NEED TO ADD AN TRYCATCH TO THIS LOOP AS WELL TO AVOID TIME OUT ERRORS, SEE THE NEXT LOOP BELOW
```{r}
cs_df$gs_id <- "" #don't make it empty everytime you run scrape! 
```


```{r}
time <- .1 # I placed the waiting time outside the loop
i <- 1 # Our loop iterator is now a variable. This means I can change it within a while loop. Using a for loop you cant change your iterator in the loop itself.
```


```{r}
while (i <= nrow(cs_df)) {
  print(i)
 
  Sys.sleep(time)

  tryCatch({
     if (cs_df[i,c("gs_id")]=="") {
       cs_df[i,c("gs_id")] <- get_scholar_id_fix(last_name = cs_df[i, c("last_name")], first_name = cs_df[i, c("first_name")],affiliation = cs_df[i,c("affiliation")])
      i <- i + 1
      time <- 1 #reset time if succesfull 
     }
    }, 
    warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
        cs_df[i,c("gs_id")] <<- NA
        i <<- i + 1
        time <<- 1 #reset time if succesfull 
        }, #BUT WE DO WANT TO CONTINUE. NOTE THE DOUBLE << THIS IS BECAUSE I WANT TO CHANGE A VARIABLE WHICH EXISTS OUTSIDE THE WARNING FUNCTION
    error=function(e){
      err <- conditionMessage(e)
      if (stringr::str_detect(err, "Google is rate limiting you")) {
        time <<- min(time + 10, 3600*2)
        cat("Error:", conditionMessage(e), "\n") #ERROR message
        cat("sleep time:", time,  "\n")
        cat("ik zit in loop", i)
      } else {
        cat("ERROR :", conditionMessage(e), "\n") # continue on error, but print the error
        i <<- i + 1
        }
      })
}
```


```{r}
# remove those without pubs from the df
# seems we're left with about 34 sociology staff members!
cs_df_copy <- cs_df #just to also have the data with staff without scholar_id
cs_df <- cs_df[!cs_df$gs_id == "", ] 
cs_df <- cs_df[!is.na(cs_df$gs_id), ] 



```


```{r}
save(cs_df, file="cs_df_s3.RData")
save(cs_df_copy, file="cs_df_copy.RData")

```

# publications and profiles

We save the publications and info of the scholar profiles in new objects. 



```{r}
soc_list_profiles <- list()  # first we create an empty list that we then fill up with the for loop
soc_list_publications <- list()
```


```{r}
time <- .1 # I placed the waiting time outside the loop
i <- 1 # Our loop iterator is now a variable. This means I can change it within a while loop. Using a for loop you cant change your iterator in the loop itself.
```


```{r}
while (i <= nrow(cs_df)) {
    print(i)
    Sys.sleep(time)

   
    tryCatch({
    #In this part of the tryCatch function you put all the stuff you want to do in the loop.
    # note how you call different elements in a list '[[]]', fill in the i-th element
    soc_list_profiles[[i]] <- get_profile(cs_df[i, c("gs_id")])  # Note how we call row i (remember how to call rows in a DF/Matrix) and then the associated scholar id
    soc_list_publications[[i]] <- get_publications(cs_df[i, c("gs_id")])
    soc_list_publications[[i]][, c("gs_id")] <- cs_df[i, c("gs_id")]  # note that we again attach an id
    # so both functions here call the entire profile and pubs for an author, based on google
    # scholar ids
    i <- i + 1 #IMPORTANT, YOU NEED TO TELL THE WHILE LOOP THAT YOUR ITERATOR HAS TO BE INCREASED
    time <- .1
    },
      warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
        i <<- i + 1
        }, #BUT WE DO WANT TO CONTINUE. NOTE THE DOUBLE << THIS IS BECAUSE I WANT TO CHANGE A VARIABLE WHICH EXISTS OUTSIDE THE WARNING FUNCTION
      error =function(e) {
        time <<- min(time + 10, 3600*2)
        cat("Error:", conditionMessage(e), "\n") #ERROR message
        cat("sleep time:", time,  "\n")
        cat("ik zit in loop", i)
        #AFTER THE NEW SLEEP TIME, WE TRY AGAIN, WE THEREFORE DO NOT UPDATE i. ideally you also want to have some break option. And maybe you also want to save your data when you hit a time out error. 
      })
   
 
}
```


```{r}
cs_df_publications <- bind_rows(soc_list_publications)
cs_list_profiles <- soc_list_profiles
```


```{r}
save(cs_df_publications, file="cs_df_publications.RData")
save(cs_list_profiles, file= "cs_list_profiles.RData")
```



## put the info of the profiles in our data set of staff members cs_df

Please note, I got an error with the original script (as in the book) so needed to tweak things a little. 
```{r}
cs_profiles_df <- list()


for (i in 1:length(cs_list_profiles)) {
    # soc_profiles_df[[i]] <- data.frame(t(unlist(soc_list_profiles[[i]][1:8]))) #some annyoing
    # data handling
  if (!is.null(cs_list_profiles[[i]])) {
    cs_profiles_df[[i]] <- unlist(cs_list_profiles[[i]][1:8])
    cs_profiles_df[[i]] <- data.frame(cs_profiles_df[[i]])
    cs_profiles_df[[i]] <- t(cs_profiles_df[[i]])
    row.names(cs_profiles_df[[i]]) <- NULL
    cs_profiles_df[[i]] <- data.frame(cs_profiles_df[[i]])
  }
}

#soc_profiles_df

cs_profiles_df2 <- bind_rows(cs_profiles_df)
cs_df <- left_join(cs_df, cs_profiles_df2, by = c(gs_id = "id"))  # merge data with cs_df
cs_df  # notice all the new information we were able to get from the scholar profiles!

```


```{r}
save(cs_df, file="cs_df_s4.RData")

```


# citation history

```{r}
# get citation history of a scholar
cs_staff_cit <- list()
```


```{r}
time <- .1 # I placed the waiting time outside the loop
i <- 1 # Our loop iterator is now a variable. This means I can change it within a while loop. Using a for loop you cant change your iterator in the loop itself.
```


```{r}
while (i <= nrow(cs_df)) {
    print(i)
    Sys.sleep(time)

    tryCatch({
      cs_staff_cit[[i]] <- get_citation_history(cs_df[i, c("gs_id")])
        if (nrow(cs_staff_cit[[i]]) > 0) {
          cs_staff_cit[[i]][, c("gs_id")] <- cs_df[i, c("gs_id")]  # again attach the gs_id as third column
        }
    i <- i + 1
    time <- .1
    },
      warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
        i <<- i + 1}, #BUT WE DO WANT TO CONTINUE. NOTE THE DOUBLE << THIS IS BECAUSE I WANT TO CHANGE A VARIABLE WHICH EXISTS OUTSIDE THE WARNING FUNCTION
      error =function(e) {
        time <<- min(time + 10, 3600*2)
        cat("Error:", conditionMessage(e), "\n") #ERROR message
        cat("sleep time:", time,  "\n")
        cat("ik zit in loop", i)
        #AFTER THE NEW SLEEP TIME, WE TRY AGAIN, WE THEREFORE DO NOT UPDATE i. ideally you also want to have some break option. And maybe you also want to save your data when you hit a time out error. 
      })

    
}
cs_staff_cit <- bind_rows(cs_staff_cit)
```


```{r}
save(cs_staff_cit, file="cs_staff_cit.RData")

```


# collaborators

PLEASE NOTE, I AM ONLY COLLECTING INFORMATION ON WHO WORKS TOGETHER WITH WHOM WITHIN THE DEPARTMENT!!!

IF YOU WANT TO COLLECT MORE COLLABORATORS, PLEASE DO SO. bUT THESE WILL NOT BE NODES IN YOUR rSIENA MODEL LATER. tHUS YOU MAY USE THESE TO CONSTRUCT EGO-LEVEL COVARIATES!! 

## fcolabs

```{r}
require(rvest)
require(xml2)
require(tidyverse)

# function to get collaborators and names from GS profiles
fcollabs <- function(gsid, lookforcollabs) {

  htmlpage1 <- read_html(paste0("https://scholar.google.nl/citations?user=", gsid, "&hl=en")) # so we paste the google scholar id
  profilename <- htmlpage1 %>% html_nodes(xpath = "//*/div[@id='gsc_prf_in']") %>% html_text() # we extract the profile name of that google scholar page
  profilecollabs1 <- as.data.frame(0) # empty df necessary for later
  profilecollabs2 <- as.data.frame(0) # empty df necessary for later

  if (lookforcollabs == 1) { # so if you want to look for collabs, set function to 1

    htmlpage2 <- read_html(paste0("https://scholar.google.com/citations?view_op=list_colleagues&hl=en&user=", gsid)) # so we paste the google scholar id
    profilecollabs1 <-  htmlpage2 %>% html_nodes(css="h3") %>% html_text() # get names
    profilecollabs1 <-  as.data.frame(profilecollabs1)

    profilecollabs2 <- htmlpage2 %>% html_nodes("a") %>% html_attr("href") # get the link
    profilecollabs2 <- profilecollabs2[seq_along(profilecollabs2) %% 2 > 0]
    profilecollabs2 <- substring(profilecollabs2, 23)

  }
  if (nrow(profilecollabs1)>1) { # if there ARE collabs

    profilecollabs1 <- as.data.frame(profilecollabs1) # we want to...
    profilecollabs2 <-  as.data.frame(profilecollabs2)
    profilecollabs1[,c("coauth_id")] <- profilecollabs2[,1]

    profilecollabs1[,c("gs_id")] <- gsid #... add gs_ids of focal GS profile
    profilecollabs1[,c("name")] <- profilename #...and the the profile name of GS profile attached

    names(profilecollabs1)[1] <- "coauth"

  } else {
    profilecollabs1 <- as.data.frame(cbind(gsid, profilename)) # if NOT looking for collabs...
    names(profilecollabs1) <- c("gs_id", "name") #...we only attach gs_id and profilename

  }
  return(profilecollabs1)

}



```

```{r}
# input a google scholar id and a 1 (if you want to find collabs) or 0 (only extracting names)

cs_collabs <- list()
```


```{r}
time <- .1 
i <- 1
```


```{r}
while (i <= nrow(cs_df)) {
    print(i)
    Sys.sleep(time)
    tryCatch({
      cs_collabs[[i]] <- fcollabs(cs_df[i, c("gs_id")], 1)
      i <- i + 1
      time <- .1
    },
      warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
        i <<- i + 1}, #BUT WE DO WANT TO CONTINUE. NOTE THE DOUBLE << THIS IS BECAUSE I WANT TO CHANGE A VARIABLE WHICH EXISTS OUTSIDE THE WARNING FUNCTION
      error =function(e) {
        time <<- min(time + 10, 3600*2)
        cat("Error:", conditionMessage(e), "\n") #ERROR message
        cat("sleep time:", time,  "\n")
        cat("ik zit in loop", i)
        #AFTER THE NEW SLEEP TIME, WE TRY AGAIN, WE THEREFORE DO NOT UPDATE i. ideally you also want to have some break option. And maybe you also want to save your data when you hit a time out error. 
      })

    
}
```


```{r}

for (i in 1:length(cs_collabs)) {
  cs_df$ncollabs[i] <- nrow(cs_collabs[[i]]) 
}

cs_collabs2 <- bind_rows(cs_collabs)  # bind rows, get the unique ones!
cs_collabs_unique <- unique(cs_collabs2[, c("gs_id")]) 
cs_collabs_unique <- cs_collabs_unique[!is.na(cs_collabs_unique)]
#save(cs_collabs, file = "addfiles/cs_df_collabs1.RData")  # you notice this takes a while, so we save the data here
```

```{r}
save(cs_collabs, file="cs_collabs.RData")
save(cs_collabs2, file="cs_collabs2.RData")
save(cs_collabs_unique, file="cs_collabs_uniques.RData")
save(cs_df, file="cs_df_s5.RData")

```

