---
title: "Datascraping"
#bibliography: references.bib
author: "Daphne Janssen"
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r Sys.time()`

<!---
Last compiled on `r format(Sys.time(), '%B, %Y')`
--->

<br>

----

# Datasrcaping Sociology staff RU

## getting started

```{r}
#start with clean workspace 
rm(list=ls())
getwd()
```

## packages
```{r}
#install.packages("data.table") 
library(data.table) 
library(tidyverse) 
#install.packages("xml2")
require(xml2)
#install.packages("rvest")
require(rvest)
#install.packages("devtools")
require(devtools)
require(scholar)
```

## save webpage with staff names 

```{r}
soc_staff <- read_html("https://www.ru.nl/sociology/research/staff/")

#head(soc_staff)
#class(soc_staff)
```

## from xml to nice df of names

### selecting the table with names

```{r}
soc_staff <- soc_staff %>%
    rvest::html_nodes("body") %>%
    xml2::xml_find_all("//td") %>%
    rvest::html_text()
```

### selecting only the odd rowes with the names
```{r}
fodd <- function(x) {
  #what is x, x is a vector
 x%%2 != 0 
}

nstaf <- length(soc_staff)

soc_names <- soc_staff[fodd(1:nstaf)] 
head(soc_names)

```

### selecting only the even rowes with the expertise
```{r}
soc_experts <- soc_staff[!fodd(1:nstaf)]
head(soc_experts)
```

### combining names and expertise in soc_df 
```{r}
soc_df <- data.frame(cbind(soc_names, soc_experts))  
```


## Cleaning the dataset  

### delete rows without name info
```{r}
delrows <- which(soc_df$soc_names == "Staff:" | soc_df$soc_names == "PhD:" | soc_df$soc_names == "External PhD:" |
    soc_df$soc_names == "Guest researchers:" | soc_df$soc_names == "Other researchers:")

soc_df <- soc_df[-delrows, ]
```

### cleaning names a bit  
```{r}
# Last name seems to be everything before the comma
soc_df$last_name <- gsub(",.*$", "", soc_df$soc_names)

# first name is everything between brackets
soc_df$first_name <- as.character(str_extract_all(soc_df$soc_names, "(?<=\\().+?(?=\\))", simplify = TRUE))
```

```{r}
soc_df$last_name <- gsub(" J. \\(Jansje\\) van MSc", "", soc_df$last_name)
soc_df$first_name <- tolower(soc_df$first_name)  # everything to lower!
soc_df$last_name <- tolower(soc_df$last_name)
```

```{r}
# trimws looses all spacing before and after (if you specify 'both') a character string
soc_df$last_name <- trimws(soc_df$last_name, which = c("both"), whitespace = "[ \t\r\n]")
soc_df$first_name <- trimws(soc_df$first_name, which = c("both"), whitespace = "[ \t\r\n]")

soc_df$first_name <- as.character(str_split(soc_df$first_name, pattern=" ", n = 2, simplify = TRUE)[,1])
soc_df$first_name <- as.character(str_split(soc_df$first_name, pattern="-", n = 2, simplify = TRUE)[,1])

#removing strange characters
soc_df$first_name <- iconv(soc_df$first_name, from = 'UTF-8', to = 'ASCII//TRANSLIT')

soc_df$soc_experts <- trimws(soc_df$soc_experts, which = c("both"), whitespace = "[ \t\r\n]")
soc_df$soc_names <- trimws(soc_df$soc_names, which = c("both"), whitespace = "[ \t\r\n]")
soc_df$first_name
```

```{r}
save(soc_df, file="soc_df_s1.RData")
```


### adding gender to dataset

#### fgender

```{r, eval=FALSE}
fgender <- function(firstname_df, me, file=NULL) {

####################################
# Author: Bas Hofstra, Anne Maaike Mulders, Jochem Tolsma
# DAte:   13-10-2021, last edit: 22-09-2022
# Tasks:  - assign gender baed on name
#         - Adapted from Rense Corten code April 2021
####################################


#Input: 
#  - firstname_df: a data.frame with a column named firstname  
#  - me: a character vector introducing yourself: e.g. "J Tolsma, Radboud University"
#  - file: location and name of file to be saved. 
  
#------------------------------------------------------------------------------------
# Load required packages

if (!require("tidyverse", character.only = TRUE)) {
  install.packages("tidyverse", dependencies = TRUE)
  library(tidyverse, character.only = TRUE)
}

if (!require("rvest", character.only = TRUE)) {
  install.packages("rvest", dependencies = TRUE)
  library(rvest, character.only = TRUE)
}

if (!require("polite", character.only = TRUE)) {
  install.packages("polite", dependencies = TRUE)
  library(polite, character.only = TRUE)
}



# make links to scrape
firstname_df$name_url <- paste0("https://www.meertens.knaw.nl/nvb/naam/is/", firstname_df[, c("firstname")])



#------------------------------------------------------------------------------------
### 2: introduce to server ###

# Introduce myself to the server
session <- bow("https://www.meertens.knaw.nl/nvb/naam/is", user_agent = me , delay = 1)


#------------------------------------------------------------------------------------
### 3: make function to get table from ###
  fnames <- function(link){ 
    name_session <-nod(session, path = link)
    name_page <- scrape(name_session) 
    return(name_page)
  }
  
name_list <- list()
table_list <- list()
firstname_df$gender <- NA

  for (i in 1:nrow(firstname_df)) {
    print(i)
    name_list[[i]] <- fnames(firstname_df[i, c("name_url")])
    # extract name frequency table and gender info
    table_list[[i]] <- name_list[[i]] %>% html_table()
    
    table_list[[i]][[1]][table_list[[i]][[1]]=="--"] <- "0"
    
    if (as.numeric(table_list[[i]][[1]]$X3[2]) > as.numeric(table_list[[i]][[1]]$X3[6])) {
      firstname_df$gender[i] <- "male" } else {
        firstname_df$gender[i] <- "female"
      }
    
    if (!is.null(file)) (save(firstname_df, file=file))
    
    }
  return(firstname_df)
}
```

#### gender to soc_df
```{r, eval=FALSE}
soc_df %>% mutate(firstname=first_name) -> soc_df

soc_df$firstname
soc_df <- fgender(soc_df, me="Jochem Tolsma, RU/RUG")
```

```{r, eval=FALSE}
save(soc_df, file="soc_df_s2.RData")
```

```{r, echo=FALSE}
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_df_s2.RData")
```


### adding some other data manually

#### affiliation
```{r}
soc_df$affiliation <- "radboud university"
```


## let harvest data from google scholar

### fixing a bug in the get_scholar_id function. 

```{r, eval=FALSE}
#require(scholar)

get_scholar_id_fix <- function (last_name = "", first_name = "", affiliation = NA)
{
  if (!any(nzchar(c(first_name, last_name))))
    stop("At least one of first and last name must be specified!")
  site <- getOption("scholar_site")
  url <- paste0(site, "/citations?view_op=search_authors&mauthors=",
                first_name, "+", last_name, "&hl=en&oi=ao")
  page <- get_scholar_resp(url)
  if (is.null(page))
    return(NA)
  aa <- httr::content(page, as = "text")
  # added by Bas Hofstra: bugfix for IDs that have a dash ("-")
  ids <- substring(aa, regexpr(";user=", aa))
  ids <- substr(ids, 1, 19) # error prone, but unsure how to solve otherwise
  # if (nchar(stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]")[[1]][1]) < 18) {
  #   ids <- stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]+[[:alnum:]]+[[:punct:]]")
  # } else {
  #   ids <- stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]")
  # }
  if (length(unlist(ids)) == 0) {
    message("No Scholar ID found.")
    return(NA)
  }
  ids <- ids %>% unlist %>% gsub(";user=|[[:punct:]]$", "",
                                 .) %>% unique
  if (length(ids) > 1) {
    profiles <- lapply(ids, scholar::get_profile)
    if (is.na(affiliation)) {
      x_profile <- profiles[[1]]
      warning("Selecting first out of ", length(profiles),
              " candidate matches.")
    }
    else {
      which_profile <- sapply(profiles, function(x) {
        stringr::str_count(string = x$affiliation, pattern = stringr::coll(affiliation,
                                                                           ignore_case = TRUE))
      })
      if (all(which_profile == 0)) {
        warning("No researcher found at the indicated affiliation.")
        return(NA)
      }
      else {
        x_profile <- profiles[[which(which_profile !=
                                       0)]]
      }
    }
  }
  else {
    x_profile <- scholar::get_profile(id = ids)
  }
  return(x_profile$id)
}
```

### scholars id. 
```{r, eval=FALSE}

soc_df$gs_id <- ""

for (i in 1:nrow(soc_df)) {
  print(i)
  time <- runif(1, 0, 1)
  Sys.sleep(time)

  tryCatch({
     soc_df[i,c("gs_id")] <- get_scholar_id_fix(last_name = soc_df[i, c("last_name")], # so search on last_name of staff (third column)
                                             first_name = soc_df[i, c("first_name")],  # search on first_name of staff (fourth column)
                                             affiliation = soc_df[i,c("affiliation")]) # search on affiliation of each staff (fifth column)

    }, error=function(e){cat("ERROR :", conditionMessage(e), "\n")}) # continue on error, but print the error
  }

# remove those without pubs from the df
soc_df_copy <- soc_df #just to also have the data with staff without scholar_id
soc_df <- soc_df[!soc_df$gs_id == "", ]

```

```{r, eval=FALSE}
save(soc_df, file="soc_df_s3.RData")
save(soc_df_copy, file="soc_df_copy.RData")
```

```{r, echo=FALSE}
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_df_s3.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_df_copy.RData")
```



## publications and profiles

```{r, eval=FALSE}
soc_list_profiles <- list()  # first we create an empty list that we then fill up with the for loop
soc_list_publications <- list()

time <- 1 # I placed the waiting time outside the loop
i <- 1 # Our loop iterator is now a variable. This means I can change it within a while loop. Using a for loop you cant change your iterator in the loop itself. 

while (i <= nrow(soc_df)) {
    print(i)
    Sys.sleep(time)

   
    tryCatch({
    #In this part of the tryCatch function you put all the stuff you want to do in the loop.
    # note how you call different elements in a list '[[]]', fill in the i-th element
    soc_list_profiles[[i]] <- get_profile(soc_df[i, c("gs_id")])  # Note how we call row i (remember how to call rows in a DF/Matrix) and then the associated scholar id
    soc_list_publications[[i]] <- get_publications(soc_df[i, c("gs_id")])
    soc_list_publications[[i]][, c("gs_id")] <- soc_df[i, c("gs_id")]  # note that we again attach an id
    # so both functions here call the entire profile and pubs for an author, based on google
    # scholar ids
    i <- i + 1 #IMPORTANT, YOU NEED TO TELL THE WHILE LOOP THAT YOUR ITERATOR HAS TO BE INCREASED
    },
      warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
        i <<- i + 1}, #BUT WE DO WANT TO CONTINUE. NOTE THE DOUBLE << THIS IS BECAUSE I WANT TO CHANGE A VARIABLE WHICH EXISTS OUTSIDE THE WARNING FUNCTION
      error =function(e) {
        time <<- min(time *10, 3600*2)
        cat("Error:", conditionMessage(e), "\n") #ERROR message
        cat("sleep time:", time,  "\n")
        cat("ik zit in loop", i)
        #AFTER THE NEW SLEEP TIME, WE TRY AGAIN, WE THEREFORE DO NOT UPDATE i. ideally you also want to have some break option. And maybe you also want to save your data when you hit a time out error. 
      })
   
 
}
```


```{r, eval=FALSE}
soc_df_publications <- bind_rows(soc_list_publications)
```


```{r, eval=FALSE}
save(soc_df_publications, file="soc_df_publications.RData")
save(soc_list_profiles, file= "soc_list_profiles.RData")
```

```{r, echo=FALSE}
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_df_publications.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_list_profiles.RData")
```


### put the info of the profiles in our data set of staff members soc_df

```{r, eval=FALSE}
soc_profiles_df <- list()


for (i in 1:length(soc_list_profiles)) {
    # soc_profiles_df[[i]] <- data.frame(t(unlist(soc_list_profiles[[i]][1:8]))) 
  if (!is.null(soc_list_profiles[[i]])) {
    soc_profiles_df[[i]] <- unlist(soc_list_profiles[[i]][1:8])
    soc_profiles_df[[i]] <- data.frame(soc_profiles_df[[i]])
    soc_profiles_df[[i]] <- t(soc_profiles_df[[i]])
    row.names(soc_profiles_df[[i]]) <- NULL
    soc_profiles_df[[i]] <- data.frame(soc_profiles_df[[i]])
  }
}

#soc_profiles_df

soc_profiles_df2 <- bind_rows(soc_profiles_df)
soc_df <- left_join(soc_df, soc_profiles_df2, by = c(gs_id = "id"))  
soc_df  
```


```{r, eval=FALSE}
save(soc_df, file="soc_df_s4.RData")
```

```{r, echo=FALSE}
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_df_s4.RData")
```

## citation history

```{r, eval=FALSE}
# get citation history of a scholar
soc_staff_cit <- list()


time <- 1 # I placed the waiting time outside the loop
i <- 1 # Our loop iterator is now a variable. This means I can change it within a while loop. Using a for loop you cant change your iterator in the loop itself. 

while (i <= nrow(soc_df)) {
    print(i)
    Sys.sleep(time)

    tryCatch({
      soc_staff_cit[[i]] <- get_citation_history(soc_df[i, c("gs_id")])
        if (nrow(soc_staff_cit[[i]]) > 0) {
          soc_staff_cit[[i]][, c("gs_id")] <- soc_df[i, c("gs_id")]  # again attach the gs_id as third column
        }
    i <- i + 1
    },
      warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
        i <<- i + 1}, #BUT WE DO WANT TO CONTINUE. NOTE THE DOUBLE << THIS IS BECAUSE I WANT TO CHANGE A VARIABLE WHICH EXISTS OUTSIDE THE WARNING FUNCTION
      error =function(e) {
        time <<- min(time *10, 3600*2)
        cat("Error:", conditionMessage(e), "\n") #ERROR message
        cat("sleep time:", time,  "\n")
        cat("ik zit in loop", i)
        #AFTER THE NEW SLEEP TIME, WE TRY AGAIN, WE THEREFORE DO NOT UPDATE i. ideally you also want to have some break option. And maybe you also want to save your data when you hit a time out error. 
      })

    
}
soc_staff_cit <- bind_rows(soc_staff_cit)
```


```{r, eval=FALSE}
save(soc_staff_cit, file="soc_staff_cit.RData")
```

```{r, echo=FALSE}
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_staff_cit.RData")
```
```

## collaborators

### fcolabs

```{r, eval=FALSE}
require(rvest)
require(xml2)
require(tidyverse)

# function to get collaborators and names from GS profiles
fcollabs <- function(gsid, lookforcollabs) {

  htmlpage1 <- read_html(paste0("https://scholar.google.nl/citations?user=", gsid, "&hl=en")) # so we paste the google scholar id
  profilename <- htmlpage1 %>% html_nodes(xpath = "//*/div[@id='gsc_prf_in']") %>% html_text() # we extract the profile name of that google scholar page
  profilecollabs1 <- as.data.frame(0) # empty df necessary for later
  profilecollabs2 <- as.data.frame(0) # empty df necessary for later

  if (lookforcollabs == 1) { # so if you want to look for collabs, set function to 1

    htmlpage2 <- read_html(paste0("https://scholar.google.com/citations?view_op=list_colleagues&hl=en&user=", gsid)) # so we paste the google scholar id
    profilecollabs1 <-  htmlpage2 %>% html_nodes(css="h3") %>% html_text() # get names
    profilecollabs1 <-  as.data.frame(profilecollabs1)

    profilecollabs2 <- htmlpage2 %>% html_nodes("a") %>% html_attr("href") # get the link
    profilecollabs2 <- profilecollabs2[seq_along(profilecollabs2) %% 2 > 0]
    profilecollabs2 <- substring(profilecollabs2, 23)

  }
  if (nrow(profilecollabs1)>1) { # if there ARE collabs

    profilecollabs1 <- as.data.frame(profilecollabs1) # we want to...
    profilecollabs2 <-  as.data.frame(profilecollabs2)
    profilecollabs1[,c("coauth_id")] <- profilecollabs2[,1]

    profilecollabs1[,c("gs_id")] <- gsid #... add gs_ids of focal GS profile
    profilecollabs1[,c("name")] <- profilename #...and the the profile name of GS profile attached

    names(profilecollabs1)[1] <- "coauth"

  } else {
    profilecollabs1 <- as.data.frame(cbind(gsid, profilename)) # if NOT looking for collabs...
    names(profilecollabs1) <- c("gs_id", "name") #...we only attach gs_id and profilename

  }
  return(profilecollabs1)

}



```

```{r, eval=FALSE}
# input a google scholar id and a 1 (if you want to find collabs) or 0 (only extracting names)

soc_collabs <- list()


time <- 1 
i <- 1 
while (i <= nrow(soc_df)) {
    print(i)
    Sys.sleep(time)
    tryCatch({
      soc_collabs[[i]] <- fcollabs(soc_df[i, c("gs_id")], 1)
      i <- i + 1
    },
      warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
        i <<- i + 1}, #BUT WE DO WANT TO CONTINUE. NOTE THE DOUBLE << THIS IS BECAUSE I WANT TO CHANGE A VARIABLE WHICH EXISTS OUTSIDE THE WARNING FUNCTION
      error =function(e) {
        time <<- min(time *10, 3600*2)
        cat("Error:", conditionMessage(e), "\n") #ERROR message
        cat("sleep time:", time,  "\n")
        cat("ik zit in loop", i)
        #AFTER THE NEW SLEEP TIME, WE TRY AGAIN, WE THEREFORE DO NOT UPDATE i. ideally you also want to have some break option. And maybe you also want to save your data when you hit a time out error. 
      })

    
}





# for (i in 1:nrow(soc_df)) {
#     print(i)
#     time <- runif(1, 0, 1)
#     Sys.sleep(time)
# }

for (i in 1:length(soc_collabs)) {
  soc_df$ncollabs[i] <- nrow(soc_collabs[[i]]) 
}

soc_collabs2 <- bind_rows(soc_collabs)  # bind rows, get the unique ones!
soc_collabs_unique <- unique(soc_collabs2[, c("gs_id")]) 
soc_collabs_unique <- soc_collabs_unique[!is.na(soc_collabs_unique)]
#save(soc_collabs, file = "addfiles/soc_df_collabs1.RData")  # you notice this takes a while, so we save the data here
```

```{r, eval=FALSE}
save(soc_collabs, file="soc_collabs.RData")
save(soc_collabs2, file="soc_collabs2.RData")
save(soc_collabs_unique, file="soc_collabs_uniques.RData")
save(soc_df, file="soc_df_s5.RData")
```

```{r, echo=FALSE}
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_collabs.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_collabs2.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_collabs_uniques.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_df_s5.RData")
```


## Scraping ethnicity

```{r, eval=FALSE}
load("soc_df_s5.RData")
soc_df_6 %>% mutate(lastname=last_name) -> soc_df_6
lastname_df <- soc_df_6
```

```{r, eval=FALSE}
# setting prefixes correctly for the scraper
voorvoegsels <- c("'t ", "d' ", "de ", "de la ", "den ", "del ", "der ", "des ", "el ", "el- ", "in 't ", "la ", "le ", "les ", "op den ", "ten ", "ter ", "tes ", "van ", "van 't ", "van de " , "van der ", "van den ", "von der ", "op den ", "ul ") 

for (i in 1: length(lastname_df$lastname)) {
  if (sum(str_detect(lastname_df$lastname[i], voorvoegsels))>0) {
    last <-  as.character(str_split(lastname_df$lastname[i], pattern=" ", simplify = TRUE))
    last <- last[length(last)]
    first <- as.character(unlist(strsplit(lastname_df$lastname[i], split=last, fixed=TRUE)))
    lastname_df$lastname[i] <- paste(last, ", ", first, sep="")
  }
}

# removing double names
for (i in 1: length(lastname_df$lastname)) {
  if (!sum(str_detect(lastname_df$lastname[i], voorvoegsels))>0) {
    lastname_df$lastname[i] <- as.character(str_split(lastname_df$lastname[i], pattern=" ", n = 2, simplify = TRUE)[,1])
    lastname_df$lastname[i] <- as.character(str_split(lastname_df$lastname[i], pattern="-", n = 2, simplify = TRUE)[,1])
  }
}

lastname_df$lastname<- trimws(lastname_df$lastname, which = c("right"), whitespace = "[ \t\r\n]")
lastname_df$lastname <- str_replace_all(lastname_df$lastname, " ", "%20") #html links houden niet van spaties. 

lastname_df$np <- ""
```
 

```{r, eval=FALSE}
# Load required packages

if (!require("tidyverse", character.only = TRUE)) {
  install.packages("tidyverse", dependencies = TRUE)
  library(tidyverse, character.only = TRUE)
}

if (!require("rvest", character.only = TRUE)) {
  install.packages("rvest", dependencies = TRUE)
  library(rvest, character.only = TRUE)
}

# if (!require("polite", character.only = TRUE)) {
#   install.packages("polite", dependencies = TRUE)
#   library(polite, character.only = TRUE)
# }
#   
# if (!require("xml2", character.only = TRUE)) {
#   install.packages("xml2", dependencies = TRUE)
#   library(polite, character.only = TRUE)
# }
```

# Creating links for the website
```{r, eval=FALSE}
# creating URLs: origin
lastname_df$name_origin <- ifelse((lastname_df$np==""),
                                paste0("https://www.cbgfamilienamen.nl/nfb/detail_naam.php?gba_naam=",
                                       stringr::str_to_title(lastname_df[, c("lastname")]), 
                                       "&gba_naam=",
                                       stringr::str_to_title(lastname_df[, c("lastname")]),
                                       "&nfd_naam=",
                                       stringr::str_to_title(lastname_df[, c("lastname")]), 
                                       "&info=analyse+en+verklaring&operator=eq&taal="), 
                                paste0("https://www.cbgfamilienamen.nl/nfb/detail_naam.php?gba_naam=",
                                       lastname_df[, c("np2")],
                                       stringr::str_to_title(lastname_df[, c("lastname")]), 
                                       "&gba_naam=",
                                       lastname_df[, c("np2")],
                                       stringr::str_to_title(lastname_df[, c("lastname")]), 
                                       "&nfd_naam=",
                                       stringr::str_to_title(lastname_df[, c("lastname")]),
                                       "%2C+",
                                       lastname_df[, c("np")],
                                       "&info=analyse+en+verklaring&operator=eq&taal="))
```

# Saving everything
```{r, eval=FALSE}
name_originl <- list()
table_originl <- list()
time <- 0.1
```

## Crucial scrape loop 
```{r, eval=FALSE}

for (i in 1:nrow(lastname_df)) {
  print(i)
  Sys.sleep(time)
  tryCatch({ 
    name_originl[[i]]  <- read_html(lastname_df[i, c("name_origin")])
    table_originl[[i]] <- name_originl[[i]] %>% html_table()
  }, 
    warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
    },
    error=function(e){
      err <- conditionMessage(e)
      cat("Error:", conditionMessage(e), "\n") #ERROR message
    }  
  )
}
```

```{r, eval=FALSE}
origin_txt <- list()
for (i in 1:length(name_originl)) {
    origin_txt [[i]] <- name_originl[[i]] %>% html_text() %>% as.character()
}

```

```{r, eval=FALSE}
# Get out the relevant origin information from the xml lists
origin_ln <- list()

for (i in 1:length(name_originl)) {
  origin_ln[[i]] <- name_originl[[i]] %>% html_nodes("div") %>% rvest::html_text()
  origin_ln[[i]] <- origin_ln[[i]][[3]]
}

# Remove mess
for (i in 1:length(origin_ln)) {
  origin_ln[[i]] <- gsub("\\t", " ", origin_ln[[i]])
  origin_ln[[i]] <- gsub("\\n", " ", origin_ln[[i]])
}

# Flatten nested structure of the origin information
#origin_ln <- rbind(flatten(origin_ln))
```

```{r, eval=FALSE}
# Detaching the names and origin info for easier data handling
origin <- unlist(origin_ln)



origin <- str_extract_all(origin, "varianten(.*?)©")

# Origin information is usually mentioned after "verklaring" or "kenmerken"
origin <- str_remove_all(origin, "varianten")
origin <- str_remove_all(origin, "CBG Bronnen")
origin <- str_remove_all(origin, "catalogus")
origin <- str_remove_all(origin, "©")


verklaring <- str_remove_all(origin, "kenmerken:(.*?)$")
kenmerken <- str_extract_all(origin, "kenmerken:(.*?)$")
kenmerken <- str_remove_all(kenmerken, "specifieke componenten:(.*?)$")
sc <- str_extract_all(origin, "specifieke componenten:(.*?)$") # Not directly relevant to us, but does mean that the name has a webpage


# Make into a neat dataframe with the names attached
verklaring <- trimws(verklaring, which = "both")
kenmerken <- trimws(kenmerken, which = "both")
sc <- trimws(sc, which = "both")
vk <- data.frame(cs_df$lastname, verklaring, kenmerken, sc)
```

## Separating names with Dutch & unknown origin
Next, we identify those names for which no additional information was found. This is important to distinguish Dutch names from names with unknown origins.

- Dutch names: no label indicating that the name is Dutch, but some other information available on name origin
- Unknown names: web page cannot be found, so origin information is empty. 

```{r, eval=FALSE}
# Identify last names that could not be found
vk <- vk %>%
  mutate(verklaring = ifelse(verklaring=="", 0, verklaring), 
         kenmerken = ifelse(kenmerken=="character(0)", 0, kenmerken),
         sc = ifelse(sc=="character(0)", 0, sc),
         no_info = nchar(verklaring) + nchar(kenmerken) + nchar(sc))

vk <- vk %>%
  mutate(no_info = ifelse(no_info==3, 1, 0), 
         verklaring = ifelse(verklaring==0, NA, verklaring),
         kenmerken = ifelse(kenmerken==0, NA, kenmerken))
# If there is no text in verklaring or kenmerken, the name could not be found in the databases. 
```


## Extracting specific origin information
There are three main ways to get information about the origin of last names:

1) Under "kenmerken", last names are assigned clickable tags. These tags include unspecified foreign name tags ("andere taal"), as well as specific foreign origins of the name ("Franse naam", "Indische naam").
-> origin1 + origin4

2) Several names have more extensively written out stories behind the name, under "verklaring". A number of names contain detailed (either country-level or regional) origins, usually in the form of "De naam [xyz] is afkomstig uit [country]". 
-> origin2 

3) Some names have origin information under "verklaring" in the form of the linguistic origins of the name. This can be country specific (e.g. Chinese name), but it can also apply to multiple countries when the language is spoken in more than 1 countries (e.g. Spanish name). 
-> origin3 

```{r, eval=FALSE}

# Step 1: extracting origin tags from kenmerken
vk <- vk %>%
  mutate(origin1 = str_extract(kenmerken, "[:upper:]([:lower:]{2,}) naam"))

# Note: sometimes multiple origins are mentioned. Currently, I only extract the first one. Otherwise, we should use str_extract_all. 



# Step 2: extracting origin info from verklaring 
vk <- vk %>%
  mutate(origin2 = ifelse(as.numeric(str_detect(verklaring, "afkomstig uit")) == 1, 
         str_remove(verklaring, ".*afkomstig uit"), NA))


# Step 3: extracting additional origin info from verklaring
vk <- vk %>%
  mutate(origin3 = str_extract(verklaring, "[:upper:]([:lower:]{2,}) (achter)?(familie)?(beroeps)?naam"))



# Finally, we clean up the origin information extracted above

# Origin1: already neat
vk$origin1 <- str_remove(vk$origin1, "Joodse naam") # can be Dutch & non-Dutch

# Origin2: messy
vk$origin2 <- str_remove(vk$origin2, "\\..*") # remove extra info in the following sentence 
vk$origin2 <- str_remove(vk$origin2, "\\;.*") # remove extra info in the following sentence 
vk$origin2 <- str_remove(vk$origin2, "\\(.*") # remove extra info in the following sentence 


vk$dpg <- as.numeric(str_detect(vk$origin2, "(dorp)|(plaats)|(gemeente)|(graafschap)|(stad)|(deel)|(Friesland)")) # origin info too regional 
vk <- vk %>% mutate(origin2 = ifelse((dpg==1), NA, origin2)) # removing regional origin info
vk <- subset(vk, select = -dpg) # removing intermediate variable

# Sometimes, there were multiple countries mentioned. Take only the first:
vk$origin2 <- str_remove(vk$origin2, "\\,.*") # Only first
vk$origin2 <- str_remove(vk$origin2, "\\s(en).*") # Only first 
vk$origin2 <- str_remove(vk$origin2, "\\s(of).*") # Only first 


# Origin3: pretty neat
vk$origin3 <- str_remove(vk$origin3, "D(i)?e(ze)? (familie)?(achter)?(beroeps)?naam") # slipped through the regex
vk$origin3 <- str_remove(vk$origin3, "Een (familie)?(achter)?(beroeps)?naam") # slipped through the regex
vk$origin3 <- str_remove(vk$origin3, "Zijn (familie)?(achter)?(beroeps)?naam") # slipped through the regex
vk$origin3 <- str_remove(vk$origin3, "Als (familie)?(achter)?(beroeps)?naam") # slipped through the regex
vk$origin3 <- str_remove(vk$origin3, "Joodse (familie)?(achter)?naam")
vk$origin3 <- str_remove(vk$origin3, "Bijbelse (familie)?(achter)?naam")


# Setting empty origin variables to NA (Dutch or unfound foreign)
vk <- vk %>%
  mutate(origin1 = as.character(ifelse(origin1==""|origin1=="character(0)", NA, origin1)),
         origin2 = as.character(ifelse(origin2==""|origin2=="character(0)", NA, origin2)),
         origin3 = as.character(ifelse(origin3==""|origin3=="character(0)", NA, origin3)))



# Finally, the tag "andere taal" was used to distinguish foreign names of unknown origin from known Dutch names. 
vk <- vk %>%
  mutate(origin4 = ifelse((as.numeric(str_detect(kenmerken, "andere taal"))==1), "non-Dutch", NA))


```

<br>

# Datasrcaping Computing science staff RU

## getting started

```{r, eval=FALSE}
#start with clean workspace 
rm(list=ls())
getwd()
```

## packages
```{r, eval=FALSE}
#install.packages("data.table") 
library(data.table) 
library(tidyverse) 
#install.packages("xml2")
require(xml2)
#install.packages("rvest")
require(rvest)
#install.packages("devtools")
require(devtools)

require(scholar)
```

## save webpage with staff names 

```{r, eval=FALSE}
cs_staff <- read_html("https://www.cs.ru.nl/staff/index.html")

#head(soc_staff)
#class(soc_staff)
```

## from xml to nice df of names

### selecting the table with names

```{r, eval=FALSE}
cs_staff <- cs_staff %>%
    rvest::html_nodes("body") %>%
    xml2::xml_find_all("//a") %>%
    rvest::html_text()
```

```{r, eval=FALSE}
Encoding(cs_staff) <- "UTF-8"
cs_staff <- iconv(cs_staff, from="UTF-8", to="LATIN1")
require(stringi)
cs_staff <- stri_trans_general(cs_staff, id = "Latin-ASCII")

check <- NA

for (i in 1:length(cs_staff)) {
  check[i] <- str_length(cs_staff[i]) > 1
}

cs_staff2 <- cs_staff[check]
cs_staff <- cs_staff2
```


### names in df
```{r, eval=FALSE}
cs_df <- data.frame(cs_staff)  
```

## Cleaning the dataset  

### cleaning names a bit  
```{r, eval=FALSE}
require(stringr)
# Last name seems to be everything after ) 
cs_df$last_name <-as.character(str_split(cs_df$cs_staff, pattern="\\)", n = 2, simplify = TRUE)[,2])

# first name is everything between brackets
cs_df$first_name <- as.character(str_extract_all(cs_df$cs_staff, "(?<=\\().+?(?=\\))", simplify = TRUE))
```

```{r, eval=FALSE}
cs_df$first_name <- tolower(cs_df$first_name)  # everything to lower!
cs_df$last_name <- tolower(cs_df$last_name)
```

```{r, eval=FALSE}
# trimws looses all spacing before and after (if you specify 'both') a character string
cs_df$last_name <- trimws(cs_df$last_name, which = c("both"), whitespace = "[ \t\r\n]")
cs_df$first_name <- trimws(cs_df$first_name, which = c("both"), whitespace = "[ \t\r\n]")

cs_df$first_name <- as.character(str_split(cs_df$first_name, pattern=" ", n = 2, simplify = TRUE)[,1])
cs_df$first_name <- as.character(str_split(cs_df$first_name, pattern="-", n = 2, simplify = TRUE)[,1])

#cs_df$last_name <- as.character(str_split(cs_df$last_name, pattern=" ", n = 2, simplify = TRUE)[,1])
#cs_df$last_name <- as.character(str_split(cs_df$last_name, pattern="-", n = 2, simplify = TRUE)[,1])


#cs_df$soc_names <- trimws(cs_df$soc_names, which = c("both"), whitespace = "[ \t\r\n]")
cs_df$last_name
```

```{r, eval=FALSE}
save(cs_df, file="cs_df_s1b.RData")
```


### adding gender to dataset

#### fgender

```{r, eval=FALSE}
fgender <- function(firstname_df, me, file=NULL) {

####################################
# Author: Bas Hofstra, Anne Maaike Mulders, Jochem Tolsma
# DAte:   13-10-2021, last edit: 22-09-2022
# Tasks:  - assign gender baed on name
#         - Adapted from Rense Corten code April 2021
####################################


#Input: 
#  - firstname_df: a data.frame with a column named firstname  and gender!
#  - me: a character vector introducing yourself: e.g. "J Tolsma, Radboud University"
#  - file: location and name of file to be saved. 
  
#------------------------------------------------------------------------------------
# Load required packages

if (!require("tidyverse", character.only = TRUE)) {
  install.packages("tidyverse", dependencies = TRUE)
  library(tidyverse, character.only = TRUE)
}

if (!require("rvest", character.only = TRUE)) {
  install.packages("rvest", dependencies = TRUE)
  library(rvest, character.only = TRUE)
}

if (!require("polite", character.only = TRUE)) {
  install.packages("polite", dependencies = TRUE)
  library(polite, character.only = TRUE)
}



# make links to scrape
firstname_df$name_url <- paste0("https://www.meertens.knaw.nl/nvb/naam/is/", firstname_df[, c("firstname")])



#------------------------------------------------------------------------------------
### 2: introduce to server ###

# Introduce myself to the server
session <- bow("https://www.meertens.knaw.nl/nvb/naam/is", user_agent = me , delay = 1)


#------------------------------------------------------------------------------------
### 3: make function to get table from ###
  fnames <- function(link){ 
    name_session <-nod(session, path = link)
    name_page <- scrape(name_session) 
    return(name_page)
  }
  
name_list <- list()
table_list <- list()


  for (i in 1:nrow(firstname_df)) {
    print(i)
    if (!(is.na(firstname_df$gender))) next
    name_list[[i]] <- fnames(firstname_df[i, c("name_url")])
    # extract name frequency table and gender info
    if (length(name_list[[i]] %>% html_table())>0) {
      
      table_list[[i]] <- name_list[[i]] %>% html_table()
      table_list[[i]][[1]][table_list[[i]][[1]]=="--"] <- "0"
      if (as.numeric(table_list[[i]][[1]]$X3[2]) > as.numeric(table_list[[i]][[1]]$X3[6])) {
        firstname_df$gender[i] <- "male" } else {
          firstname_df$gender[i] <- "female"
        }
    }
    if (!is.null(file)) (save(firstname_df, file=file))
    
    }
  return(firstname_df)
}
```

#### gender to cs_df
```{r, eval=FALSE}
cs_df %>% mutate(firstname=first_name) -> cs_df
cs_df$gender <- NA
```

```{r, eval=FALSE}
cs_df$firstname
cs_df <- fgender(cs_df, me="Jochem Tolsma, RU/RUG", file="tempgender_cs_d2.RData")

```

```{r, eval=FALSE}
save(cs_df, file="cs_df_s2b.RData") #genderized last. 
```

### adding some other data manually

#### affiliation
```{r, eval=FALSE}
cs_df$affiliation <- "radboud university"  
```

#### other stuff

```{r, eval=FALSE}
#require(scholar)

get_scholar_id_fix <- function (last_name = "", first_name = "", affiliation = NA)
{
  if (!any(nzchar(c(first_name, last_name))))
    stop("At least one of first and last name must be specified!")
  site <- getOption("scholar_site")
  url <- paste0(site, "/citations?view_op=search_authors&mauthors=",
                first_name, "+", last_name, "&hl=en&oi=ao")
  page <- get_scholar_resp(url)
  if (is.null(page))
    return(NA)
  aa <- httr::content(page, as = "text")
  # added by Bas Hofstra: bugfix for IDs that have a dash ("-")
  ids <- substring(aa, regexpr(";user=", aa))
  ids <- substr(ids, 1, 19) # error prone, but unsure how to solve otherwise
  # if (nchar(stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]")[[1]][1]) < 18) {
  #   ids <- stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]+[[:alnum:]]+[[:punct:]]")
  # } else {
  #   ids <- stringr::str_extract_all(string = aa, pattern = ";user=[[:alnum:]]+[[:punct:]]")
  # }
  if (length(unlist(ids)) == 0) {
    message("No Scholar ID found.")
    return(NA)
  }
  ids <- ids %>% unlist %>% gsub(";user=|[[:punct:]]$", "",
                                 .) %>% unique
  if (length(ids) > 1) {
    profiles <- lapply(ids, scholar::get_profile)
    if (is.na(affiliation)) {
      x_profile <- profiles[[1]]
      warning("Selecting first out of ", length(profiles),
              " candidate matches.")
    }
    else {
      which_profile <- sapply(profiles, function(x) {
        stringr::str_count(string = x$affiliation, pattern = stringr::coll(affiliation,
                                                                           ignore_case = TRUE))
      })
      if (all(which_profile == 0)) {
        warning("No researcher found at the indicated affiliation.")
        return(NA)
      }
      else {
        x_profile <- profiles[[which(which_profile !=
                                       0)]]
      }
    }
  }
  else {
    x_profile <- scholar::get_profile(id = ids)
  }
  return(x_profile$id)
}
```

### scholars id. 

```{r, eval=FALSE}
cs_df$gs_id <- "" #don't make it empty everytime you run scrape! 
```

```{r, eval=FALSE}
time <- .1 # I placed the waiting time outside the loop
i <- 1 # Our loop iterator is now a variable. This means I can change it within a while loop. Using a for loop you cant change your iterator in the loop itself.
```

```{r, eval=FALSE}
while (i <= nrow(cs_df)) {
  print(i)
 
  Sys.sleep(time)

  tryCatch({
     if (cs_df[i,c("gs_id")]=="") {
       cs_df[i,c("gs_id")] <- get_scholar_id_fix(last_name = cs_df[i, c("last_name")], first_name = cs_df[i, c("first_name")],affiliation = cs_df[i,c("affiliation")])
      i <- i + 1
      time <- 1 #reset time if succesfull 
     }
    }, 
    warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
        cs_df[i,c("gs_id")] <<- NA
        i <<- i + 1
        time <<- 1 #reset time if succesfull 
        }, #BUT WE DO WANT TO CONTINUE. NOTE THE DOUBLE << THIS IS BECAUSE I WANT TO CHANGE A VARIABLE WHICH EXISTS OUTSIDE THE WARNING FUNCTION
    error=function(e){
      err <- conditionMessage(e)
      if (stringr::str_detect(err, "Google is rate limiting you")) {
        time <<- min(time + 10, 3600*2)
        cat("Error:", conditionMessage(e), "\n") #ERROR message
        cat("sleep time:", time,  "\n")
        cat("ik zit in loop", i)
      } else {
        cat("ERROR :", conditionMessage(e), "\n") # continue on error, but print the error
        i <<- i + 1
        }
      })
}
```

```{r, eval=FALSE}
# remove those without pubs from the df
cs_df_copy <- cs_df #just to also have the data with staff without scholar_id
cs_df <- cs_df[!cs_df$gs_id == "", ] 
cs_df <- cs_df[!is.na(cs_df$gs_id), ] 
```

```{r, eval=FALSE}
save(cs_df, file="cs_df_s3.RData")
save(cs_df_copy, file="cs_df_copy.RData")
```

## publications and profiles

```{r, eval=FALSE}
# Saving the publications and info of the scholar profiles in new objects. 
soc_list_profiles <- list()  # first we create an empty list that we then fill up with the for loop
soc_list_publications <- list()
```

```{r, eval=FALSE}
time <- .1 # I placed the waiting time outside the loop
i <- 1 # Our loop iterator is now a variable. This means I can change it within a while loop. Using a for loop you cant change your iterator in the loop itself.
```

```{r, eval=FALSE}
while (i <= nrow(cs_df)) {
    print(i)
    Sys.sleep(time)

   
    tryCatch({
    #In this part of the tryCatch function you put all the stuff you want to do in the loop.
    # note how you call different elements in a list '[[]]', fill in the i-th element
    soc_list_profiles[[i]] <- get_profile(cs_df[i, c("gs_id")])  # Note how we call row i (remember how to call rows in a DF/Matrix) and then the associated scholar id
    soc_list_publications[[i]] <- get_publications(cs_df[i, c("gs_id")])
    soc_list_publications[[i]][, c("gs_id")] <- cs_df[i, c("gs_id")]  # note that we again attach an id
    # so both functions here call the entire profile and pubs for an author, based on google
    # scholar ids
    i <- i + 1 #IMPORTANT, YOU NEED TO TELL THE WHILE LOOP THAT YOUR ITERATOR HAS TO BE INCREASED
    time <- .1
    },
      warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
        i <<- i + 1
        }, #BUT WE DO WANT TO CONTINUE. NOTE THE DOUBLE << THIS IS BECAUSE I WANT TO CHANGE A VARIABLE WHICH EXISTS OUTSIDE THE WARNING FUNCTION
      error =function(e) {
        time <<- min(time + 10, 3600*2)
        cat("Error:", conditionMessage(e), "\n") #ERROR message
        cat("sleep time:", time,  "\n")
        cat("ik zit in loop", i)
        #AFTER THE NEW SLEEP TIME, WE TRY AGAIN, WE THEREFORE DO NOT UPDATE i. ideally you also want to have some break option. And maybe you also want to save your data when you hit a time out error. 
      })
   
 
}
```

```{r, eval=FALSE}
cs_df_publications <- bind_rows(soc_list_publications)
cs_list_profiles <- soc_list_profiles
```

```{r, eval=FALSE}
save(cs_df_publications, file="cs_df_publications.RData")
save(cs_list_profiles, file= "cs_list_profiles.RData")
```


### put the info of the profiles in our data set of staff members cs_df

```{r, eval=FALSE}
cs_profiles_df <- list()


for (i in 1:length(cs_list_profiles)) {
    # soc_profiles_df[[i]] <- data.frame(t(unlist(soc_list_profiles[[i]][1:8]))) #some annyoing
    # data handling
  if (!is.null(cs_list_profiles[[i]])) {
    cs_profiles_df[[i]] <- unlist(cs_list_profiles[[i]][1:8])
    cs_profiles_df[[i]] <- data.frame(cs_profiles_df[[i]])
    cs_profiles_df[[i]] <- t(cs_profiles_df[[i]])
    row.names(cs_profiles_df[[i]]) <- NULL
    cs_profiles_df[[i]] <- data.frame(cs_profiles_df[[i]])
  }
}

#soc_profiles_df

cs_profiles_df2 <- bind_rows(cs_profiles_df)
cs_df <- left_join(cs_df, cs_profiles_df2, by = c(gs_id = "id"))  
cs_df  
```

```{r, eval=FALSE}
save(cs_df, file="cs_df_s4.RData")
```


## citation history

```{r, eval=FALSE}
cs_staff_cit <- list()
```

```{r, eval=FALSE}
time <- .1 # I placed the waiting time outside the loop
i <- 1 # Our loop iterator is now a variable. This means I can change it within a while loop. Using a for loop you cant change your iterator in the loop itself.
```

```{r, eval=FALSE}
while (i <= nrow(cs_df)) {
    print(i)
    Sys.sleep(time)

    tryCatch({
      cs_staff_cit[[i]] <- get_citation_history(cs_df[i, c("gs_id")])
        if (nrow(cs_staff_cit[[i]]) > 0) {
          cs_staff_cit[[i]][, c("gs_id")] <- cs_df[i, c("gs_id")]  # again attach the gs_id as third column
        }
    i <- i + 1
    time <- .1
    },
      warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
        i <<- i + 1}, #BUT WE DO WANT TO CONTINUE. NOTE THE DOUBLE << THIS IS BECAUSE I WANT TO CHANGE A VARIABLE WHICH EXISTS OUTSIDE THE WARNING FUNCTION
      error =function(e) {
        time <<- min(time + 10, 3600*2)
        cat("Error:", conditionMessage(e), "\n") #ERROR message
        cat("sleep time:", time,  "\n")
        cat("ik zit in loop", i)
        #AFTER THE NEW SLEEP TIME, WE TRY AGAIN, WE THEREFORE DO NOT UPDATE i. ideally you also want to have some break option. And maybe you also want to save your data when you hit a time out error. 
      })

    
}
cs_staff_cit <- bind_rows(cs_staff_cit)
```

```{r, eval=FALSE}
save(cs_staff_cit, file="cs_staff_cit.RData")
```

## collaborators

### fcolabs

```{r, eval=FALSE}
require(rvest)
require(xml2)
require(tidyverse)

# function to get collaborators and names from GS profiles
fcollabs <- function(gsid, lookforcollabs) {

  htmlpage1 <- read_html(paste0("https://scholar.google.nl/citations?user=", gsid, "&hl=en")) # so we paste the google scholar id
  profilename <- htmlpage1 %>% html_nodes(xpath = "//*/div[@id='gsc_prf_in']") %>% html_text() # we extract the profile name of that google scholar page
  profilecollabs1 <- as.data.frame(0) # empty df necessary for later
  profilecollabs2 <- as.data.frame(0) # empty df necessary for later

  if (lookforcollabs == 1) { # so if you want to look for collabs, set function to 1

    htmlpage2 <- read_html(paste0("https://scholar.google.com/citations?view_op=list_colleagues&hl=en&user=", gsid)) # so we paste the google scholar id
    profilecollabs1 <-  htmlpage2 %>% html_nodes(css="h3") %>% html_text() # get names
    profilecollabs1 <-  as.data.frame(profilecollabs1)

    profilecollabs2 <- htmlpage2 %>% html_nodes("a") %>% html_attr("href") # get the link
    profilecollabs2 <- profilecollabs2[seq_along(profilecollabs2) %% 2 > 0]
    profilecollabs2 <- substring(profilecollabs2, 23)

  }
  if (nrow(profilecollabs1)>1) { # if there ARE collabs

    profilecollabs1 <- as.data.frame(profilecollabs1) # we want to...
    profilecollabs2 <-  as.data.frame(profilecollabs2)
    profilecollabs1[,c("coauth_id")] <- profilecollabs2[,1]

    profilecollabs1[,c("gs_id")] <- gsid #... add gs_ids of focal GS profile
    profilecollabs1[,c("name")] <- profilename #...and the the profile name of GS profile attached

    names(profilecollabs1)[1] <- "coauth"

  } else {
    profilecollabs1 <- as.data.frame(cbind(gsid, profilename)) # if NOT looking for collabs...
    names(profilecollabs1) <- c("gs_id", "name") #...we only attach gs_id and profilename

  }
  return(profilecollabs1)

}
```

```{r, eval=FALSE}
# input a google scholar id and a 1 (if you want to find collabs) or 0 (only extracting names)

cs_collabs <- list()
```

```{r, eval=FALSE}
time <- .1 
i <- 1
```

```{r, eval=FALSE}
while (i <= nrow(cs_df)) {
    print(i)
    Sys.sleep(time)
    tryCatch({
      cs_collabs[[i]] <- fcollabs(cs_df[i, c("gs_id")], 1)
      i <- i + 1
      time <- .1
    },
      warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
        i <<- i + 1}, #BUT WE DO WANT TO CONTINUE. NOTE THE DOUBLE << THIS IS BECAUSE I WANT TO CHANGE A VARIABLE WHICH EXISTS OUTSIDE THE WARNING FUNCTION
      error =function(e) {
        time <<- min(time + 10, 3600*2)
        cat("Error:", conditionMessage(e), "\n") #ERROR message
        cat("sleep time:", time,  "\n")
        cat("ik zit in loop", i)
        #AFTER THE NEW SLEEP TIME, WE TRY AGAIN, WE THEREFORE DO NOT UPDATE i. ideally you also want to have some break option. And maybe you also want to save your data when you hit a time out error. 
      })

    
}
```

```{r, eval=FALSE}

for (i in 1:length(cs_collabs)) {
  cs_df$ncollabs[i] <- nrow(cs_collabs[[i]]) 
}

cs_collabs2 <- bind_rows(cs_collabs)  # bind rows, get the unique ones!
cs_collabs_unique <- unique(cs_collabs2[, c("gs_id")]) 
cs_collabs_unique <- cs_collabs_unique[!is.na(cs_collabs_unique)]
#save(cs_collabs, file = "addfiles/cs_df_collabs1.RData")  
```

```{r, eval=FALSE}
save(cs_collabs, file="cs_collabs.RData")
save(cs_collabs2, file="cs_collabs2.RData")
save(cs_collabs_unique, file="cs_collabs_uniques.RData")
save(cs_df, file="cs_df_s5.RData")
```


## Scraping ethnicity

```{r, eval=FALSE}
load("cs_df_s1b.RData")
cs_df %>% mutate(lastname=last_name) -> cs_df
lastname_df <- cs_df
```

```{r, eval=FALSE}
# Setting prefixes correctly for scraper
voorvoegsels <- c("'t ", "d' ", "de ", "de la ", "den ", "del ", "der ", "des ", "el ", "el- ", "in 't ", "la ", "le ", "les ", "op den ", "ten ", "ter ", "tes ", "van ", "van 't ", "van de " , "van der ", "van den ", "von der ", "op den ", "ul ") 

for (i in 1: length(lastname_df$lastname)) {
  if (sum(str_detect(lastname_df$lastname[i], voorvoegsels))>0) {
    last <-  as.character(str_split(lastname_df$lastname[i], pattern=" ", simplify = TRUE))
    last <- last[length(last)]
    first <- as.character(unlist(strsplit(lastname_df$lastname[i], split=last, fixed=TRUE)))
    lastname_df$lastname[i] <- paste(last, ", ", first, sep="")
  }
}

# Removing duplicate names
for (i in 1: length(lastname_df$lastname)) {
  if (!sum(str_detect(lastname_df$lastname[i], voorvoegsels))>0) {
    lastname_df$lastname[i] <- as.character(str_split(lastname_df$lastname[i], pattern=" ", n = 2, simplify = TRUE)[,1])
    lastname_df$lastname[i] <- as.character(str_split(lastname_df$lastname[i], pattern="-", n = 2, simplify = TRUE)[,1])
  }
}

lastname_df$lastname<- trimws(lastname_df$lastname, which = c("right"), whitespace = "[ \t\r\n]")
lastname_df$lastname <- str_replace_all(lastname_df$lastname, " ", "%20") #html links houden niet van spaties. 

lastname_df$np <- ""
```
 

```{r, eval=FALSE}
# Load required packages

if (!require("tidyverse", character.only = TRUE)) {
  install.packages("tidyverse", dependencies = TRUE)
  library(tidyverse, character.only = TRUE)
}

if (!require("rvest", character.only = TRUE)) {
  install.packages("rvest", dependencies = TRUE)
  library(rvest, character.only = TRUE)
}

# if (!require("polite", character.only = TRUE)) {
#   install.packages("polite", dependencies = TRUE)
#   library(polite, character.only = TRUE)
# }
#   
# if (!require("xml2", character.only = TRUE)) {
#   install.packages("xml2", dependencies = TRUE)
#   library(polite, character.only = TRUE)
# }
```

```{r, eval=FALSE}
# creating URLs: origin
lastname_df$name_origin <- ifelse((lastname_df$np==""),
                                paste0("https://www.cbgfamilienamen.nl/nfb/detail_naam.php?gba_naam=",
                                       stringr::str_to_title(lastname_df[, c("lastname")]), 
                                       "&gba_naam=",
                                       stringr::str_to_title(lastname_df[, c("lastname")]),
                                       "&nfd_naam=",
                                       stringr::str_to_title(lastname_df[, c("lastname")]), 
                                       "&info=analyse+en+verklaring&operator=eq&taal="), 
                                paste0("https://www.cbgfamilienamen.nl/nfb/detail_naam.php?gba_naam=",
                                       lastname_df[, c("np2")],
                                       stringr::str_to_title(lastname_df[, c("lastname")]), 
                                       "&gba_naam=",
                                       lastname_df[, c("np2")],
                                       stringr::str_to_title(lastname_df[, c("lastname")]), 
                                       "&nfd_naam=",
                                       stringr::str_to_title(lastname_df[, c("lastname")]),
                                       "%2C+",
                                       lastname_df[, c("np")],
                                       "&info=analyse+en+verklaring&operator=eq&taal="))
```

```{r, eval=FALSE}
# Saving everything
name_originl <- list()
table_originl <- list()
time <- 0.1
```

## crucical scrape loop
```{r, eval=FALSE}
for (i in 1:nrow(lastname_df)) {
  print(i)
  Sys.sleep(time)
  tryCatch({ 
    name_originl[[i]]  <- read_html(lastname_df[i, c("name_origin")])
    table_originl[[i]] <- name_originl[[i]] %>% html_table()
  }, 
    warning = function(w) {
        cat("WARNING:", conditionMessage(w), "\n") #WARNING message
    },
    error=function(e){
      err <- conditionMessage(e)
      cat("Error:", conditionMessage(e), "\n") #ERROR message
    }  
  )
}
```

```{r, eval=FALSE}
origin_txt <- list()
for (i in 1:length(name_originl)) {
    origin_txt [[i]] <- name_originl[[i]] %>% html_text() %>% as.character()
}
```

```{r, eval=FALSE}
# Get out the relevant origin information from the xml lists
origin_ln <- list()

for (i in 1:length(name_originl)) {
  origin_ln[[i]] <- name_originl[[i]] %>% html_nodes("div") %>% rvest::html_text()
  origin_ln[[i]] <- origin_ln[[i]][[3]]
}

# Remove mess
for (i in 1:length(origin_ln)) {
  origin_ln[[i]] <- gsub("\\t", " ", origin_ln[[i]])
  origin_ln[[i]] <- gsub("\\n", " ", origin_ln[[i]])
}

# Flatten nested structure of the origin information
#origin_ln <- rbind(flatten(origin_ln))
```


```{r, eval=FALSE}
# Detaching the names and origin info for easier data handling
origin <- unlist(origin_ln)



origin <- str_extract_all(origin, "varianten(.*?)©")

# Origin information is usually mentioned after "verklaring" or "kenmerken"
origin <- str_remove_all(origin, "varianten")
origin <- str_remove_all(origin, "CBG Bronnen")
origin <- str_remove_all(origin, "catalogus")
origin <- str_remove_all(origin, "©")


verklaring <- str_remove_all(origin, "kenmerken:(.*?)$")
kenmerken <- str_extract_all(origin, "kenmerken:(.*?)$")
kenmerken <- str_remove_all(kenmerken, "specifieke componenten:(.*?)$")
sc <- str_extract_all(origin, "specifieke componenten:(.*?)$") # Not directly relevant to us, but does mean that the name has a webpage


# Make into a neat dataframe with the names attached
verklaring <- trimws(verklaring, which = "both")
kenmerken <- trimws(kenmerken, which = "both")
sc <- trimws(sc, which = "both")
vk <- data.frame(cs_df$lastname, verklaring, kenmerken, sc)
```

## Separating names with Dutch & unknown origin
Next, we identify those names for which no additional information was found. This is important to distinguish Dutch names from names with unknown origins.

- Dutch names: no label indicating that the name is Dutch, but some other information available on name origin
- Unknown names: web page cannot be found, so origin information is empty. 

```{r, eval=FALSE}

# Identify last names that could not be found
vk <- vk %>%
  mutate(verklaring = ifelse(verklaring=="", 0, verklaring), 
         kenmerken = ifelse(kenmerken=="character(0)", 0, kenmerken),
         sc = ifelse(sc=="character(0)", 0, sc),
         no_info = nchar(verklaring) + nchar(kenmerken) + nchar(sc))

vk <- vk %>%
  mutate(no_info = ifelse(no_info==3, 1, 0), 
         verklaring = ifelse(verklaring==0, NA, verklaring),
         kenmerken = ifelse(kenmerken==0, NA, kenmerken))
# If there is no text in verklaring or kenmerken, the name could not be found in the databases. 
```


## Extracting specific origin information
There are three main ways to get information about the origin of last names:

1) Under "kenmerken", last names are assigned clickable tags. These tags include unspecified foreign name tags ("andere taal"), as well as specific foreign origins of the name ("Franse naam", "Indische naam").
-> origin1 + origin4

2) Several names have more extensively written out stories behind the name, under "verklaring". A number of names contain detailed (either country-level or regional) origins, usually in the form of "De naam [xyz] is afkomstig uit [country]". 
-> origin2 

3) Some names have origin information under "verklaring" in the form of the linguistic origins of the name. This can be country specific (e.g. Chinese name), but it can also apply to multiple countries when the language is spoken in more than 1 countries (e.g. Spanish name). 
-> origin3 

```{r, eval=FALSE}

# Step 1: extracting origin tags from kenmerken
vk <- vk %>%
  mutate(origin1 = str_extract(kenmerken, "[:upper:]([:lower:]{2,}) naam"))

# Note: sometimes multiple origins are mentioned. Currently, I only extract the first one. Otherwise, we should use str_extract_all. 



# Step 2: extracting origin info from verklaring 
vk <- vk %>%
  mutate(origin2 = ifelse(as.numeric(str_detect(verklaring, "afkomstig uit")) == 1, 
         str_remove(verklaring, ".*afkomstig uit"), NA))


# Step 3: extracting additional origin info from verklaring
vk <- vk %>%
  mutate(origin3 = str_extract(verklaring, "[:upper:]([:lower:]{2,}) (achter)?(familie)?(beroeps)?naam"))



# Finally, we clean up the origin information extracted above

# Origin1: already neat
vk$origin1 <- str_remove(vk$origin1, "Joodse naam") # can be Dutch & non-Dutch

# Origin2: messy
vk$origin2 <- str_remove(vk$origin2, "\\..*") # remove extra info in the following sentence 
vk$origin2 <- str_remove(vk$origin2, "\\;.*") # remove extra info in the following sentence 
vk$origin2 <- str_remove(vk$origin2, "\\(.*") # remove extra info in the following sentence 


vk$dpg <- as.numeric(str_detect(vk$origin2, "(dorp)|(plaats)|(gemeente)|(graafschap)|(stad)|(deel)|(Friesland)")) # origin info too regional 
vk <- vk %>% mutate(origin2 = ifelse((dpg==1), NA, origin2)) # removing regional origin info
vk <- subset(vk, select = -dpg) # removing intermediate variable

# Sometimes, there were multiple countries mentioned. Take only the first:
vk$origin2 <- str_remove(vk$origin2, "\\,.*") # Only first
vk$origin2 <- str_remove(vk$origin2, "\\s(en).*") # Only first 
vk$origin2 <- str_remove(vk$origin2, "\\s(of).*") # Only first 


# Origin3: pretty neat
vk$origin3 <- str_remove(vk$origin3, "D(i)?e(ze)? (familie)?(achter)?(beroeps)?naam") # slipped through the regex
vk$origin3 <- str_remove(vk$origin3, "Een (familie)?(achter)?(beroeps)?naam") # slipped through the regex
vk$origin3 <- str_remove(vk$origin3, "Zijn (familie)?(achter)?(beroeps)?naam") # slipped through the regex
vk$origin3 <- str_remove(vk$origin3, "Als (familie)?(achter)?(beroeps)?naam") # slipped through the regex
vk$origin3 <- str_remove(vk$origin3, "Joodse (familie)?(achter)?naam")
vk$origin3 <- str_remove(vk$origin3, "Bijbelse (familie)?(achter)?naam")


# Setting empty origin variables to NA (Dutch or unfound foreign)
vk <- vk %>%
  mutate(origin1 = as.character(ifelse(origin1==""|origin1=="character(0)", NA, origin1)),
         origin2 = as.character(ifelse(origin2==""|origin2=="character(0)", NA, origin2)),
         origin3 = as.character(ifelse(origin3==""|origin3=="character(0)", NA, origin3)))



# Finally, the tag "andere taal" was used to distinguish foreign names of unknown origin from known Dutch names. 
vk <- vk %>%
  mutate(origin4 = ifelse((as.numeric(str_detect(kenmerken, "andere taal"))==1), "non-Dutch", NA))
```

