---
title: "Data"
#bibliography: references.bib
author: "Daphne Janssen"
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r Sys.time()`

<!---
Last compiled on `r format(Sys.time(), '%B, %Y')`
--->

<br>

----

# Data 

<br>

## Data collection

The data has been collected around the beginning of October 2022, through web-scraping of the Internet pages with the Sociology and Computing Science staff members. The data collected were the titles and first and last names of the individuals. Then, using an r-script, the *google scholar ids* of all staff members were scraped from the Internet. With these ids, information was obtained about the publications of the individuals, and thus the scholars with whom they published and how often they were cited. 

Around the same time, information on the *sex* of the individuals was collected through an r-script that entered the first names of all staff members into the Meertens database (https://www.meertens.knaw.nl/nvb/). Based on the sex of the highest number of individuals bearing the same name, staff members were assigned a sex. Information on the *ethnicity* of staff members was collected in a similar manner. The surnames were entered into the Dutch Family Names Database (https://www.cbgfamilienamen.nl/nfb/). Depending on whether the surname appears in this database, the staff members were assigned a Dutch or non-Dutch ethnicity. The *age* of the individuals is represented by the first year of publication. 

<br>

## Data description

### Network 

**Sample**

The staff of the department of Sociology, representing the social sciences. And the staff of the Computing Science department, representing the scientific field of STEM.

**Nodes**

The scientists within the aforementioned departments.

**Edges**

The co-authorships of scientists within (1) the Sociology department and (2) the Computing Science department. 

<br>

### Variables 

**Independent variable**

Gender: 

**Dependent variable**

Network structure: *co-authorship connections among men and women* at the diad level within the departments of Sociology and Computing Science.


**Control variables**

- The *ethnicity* of the staff members of Sociology and Computing Science.

- The *age* of the staff members of Sociology and Computing Science.

<br>

## Methods

To answer the first research question, the degree of sex segregation will be determined for both studies by calculating the density between groups and within groups for each wave. The density will also be calculated separately for the group of men and the group of women within each study. These different densities will be compared to determine whether men are more likely to work with men and/or women more likely to work with women, for both the Sociology sample and the Computer Science sample.   

To answer the second research question, an R Siena model will be estimated that includes only ego-level gender effects. This model can be used to compare the difference in the application of the gender homophily principle to the publication network of male and female scientists, between Sociology and Computer Science across the three waves.

The third research question will be answered by estimating an R Siena model that includes both ego-level gender effects and structural effects. To control for several individual characteristics, another model will be estimated that also includes covariate effects. With these models, the extent to which structural network effects affect the difference in gender homophily in the publication network of male and female scientists can be compared between Sociology and Computer Science across the three waves. 

