---
title: "Data"
#bibliography: references.bib
author: "Daphne Janssen"
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r Sys.time()`

<!---
Last compiled on `r format(Sys.time(), '%B, %Y')`
--->

<br>

----

# Data 

**Data collection**

The data has been collected around the beginning of October 2022, through web-scraping of the Internet pages with the Sociology and Computing Science staff members. The data collected were the titles and first and last names of the individuals. Then, using an r-script, the *google scholar ids* of all staff members were scraped from the Internet. With these ids, information was obtained about the publications of the individuals, and thus the scholars with whom they published and how often they were cited. 

Around the same time, information on the *sex* of the individuals was collected through an r-script that entered the first names of all staff members into the Meertens database (https://www.meertens.knaw.nl/nvb/). Based on the sex of the highest number of individuals bearing the same name, staff members were assigned a sex. Information on the *ethnicity* of staff members was collected in a similar manner. The surnames were entered into the Dutch Family Names Database (https://www.cbgfamilienamen.nl/nfb/). Depending on whether the surname appears in this database, the staff members were assigned a Dutch or non-Dutch ethnicity. The *age* of the individuals is represented by the first year of publication. 


**Sample**

The staff of the department of Sociology, representing the social sciences. And the staff of the Computing Science department, representing the scientific field of STEM.

**Nodes**

The scientists within the aforementioned departments.

**Edges**

The co-authorships of scientists within (1) the Sociology department and (2) the Computing Science department. 

<br>

**Independent variable**

-	Network structure: *co-authorship connections among men and women* at the diad level within the departments of Sociology and Computing Science.

**Dependent variable**

-	*Number of published articles*: do men who have connections with other men influence each other by publishing more articles (together) (*influence processes*). This could widen the gap between the prominence of male and female scientists.

-	*Number of citations*: do men who have connections with other men influence each other by both being more often citated (*influence processes*). This could widen the gap between the prominence of male and female scientists.

**Control variables**

-	*Number of men and women within the departments of Sociology and Computing Science.*

-	*Number of different topics men and women research within the departments of Sociology and Computing Science.*

- *The ethnicity of the staff members of Sociology and Computing Science.*

- *The age of the staff members of Sociology and Computing Science.*

