---
title: "Introduction and Research Question"
#bibliography: references.bib
author: "Daphne Janssen"
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r Sys.time()`

<!---
Last compiled on `r format(Sys.time(), '%B, %Y')`
--->

<br>

----

# Introduction   

For a long time, up to the present day, women have been underrepresented in STEM (science, technology, engineering and mathematics) studies and professions. According to the website *stemwomen.com*, the gap between men and women taking STEM courses is widest in computer sciences and engingeering & technology: 19 percent of students are women, compared to 81 percent who are men. Moreover, research shows that men account for 80 percent of the total amount of scientific production within engineering. Although women publish their papers in higher impact journals, they get less citated by the scientific engineering community compared to men (Ghiasi, Larivière & Sugimoto,2015). Furthermore, it appears that men are more likely to be authors of prestigious journals in the STEM field, compared to women (Holman, Stuart-Fox & Hauser, 2018). Also, Holman et al. (2018) estimated that STEM journals invite men to submit papers about twice as often, compared to women. These findings show that the STEM gender gap will not narrow on its own. To learn more about some specific underlying factors of the current gender gap, with the aim of contributing to a decrease in gender inequality within STEM, this paper will focus on the impact of social network characteristics of men and women researchers within the field of STEM. Previous research shows that networks are a central factor in gender inequality in obtaining leading scientific positions (Kegen, 2013). More specifically, I will examine to what extent co-authorship ties among men, and among women, but as well as between men and women, impact in gender inequality within the field of STEM. 

Previous research shows that co-authorship is highly relevant to promotion and productivity (Collins & Steffen-Fluhr, 2019). However, it is known that there are gender differences in scientific collaborations, putting women at a disadvantage (Jadidi, Karimi, Lietz, & Wagner, 2018). The gender homophily principle plays an important role in the construction of these networks. However, the literature is inconsistent on whether this principle is more common among men than women (or vice versa). While some studies find that because of the homophily principle, men in the field of STEM are more likely to collaborate with other men than with women, thus reducing the chances for women to be signalled (Collins & Steffen-Fluhr, 2019), other studies conclude that women practice more gender homophily (Jadidi et al., 2018). Ghiasi et al. (2015) find that engineers predominantly collaborate with men, contributing to scientific structures recognised by male dominance. Jadidi et al. (2018) further find that women are generally less likely to enter into collaborations that will lead to success and are more likely to be embedded in ego networks with structural holes. 

A greater number of co-authorships will likely lead to a greater number of published articles. This is another important aspect of gender inequality within STEM, but also more broadly within the scientific community. Previous research shows that women publish less scientific articles than men (Mathews & Andersen, 2001) and that women scientists contribute less to journal articles compared to the proportion they make up of the academic staff (Bird, 2011). This is problematic because academic publications are closely related to professional recognition and are an indication of a scientist’s scientific contribution (Mathews & Andersen, 2001). 
	
By looking into the differences in co-authorship patterns among men and women within the social sciences and the scientific field of STEM, it becomes possible to investigate the role of ties in gender inequality within science. A social network perspective is of crucial value to thoroughly investigate the role of ties, more specifically the severeness of the homophily principle, for the academic gender gap. This perspective allows for investigating the way the specific manifestations of gender inequality in ties works their way to the surface through ties.I would like to contribute to the current scientific field concerning gender differences in social networks within STEM, by investigating to what extent men and women differ in who they collaborate with, and to what extent this differs between sociology and engineering within three Dutch universities. Based on the above I derived the following **research question(s)**:
  
*To what extent is there a difference between men and women at Radboud University in who they work with, within Sociology and Communication science, and Engineering and Computing science?*

*And to what extent does this result in gender inequality in terms of the number of articles published and the number of citations per article, between men and women at Radboud University within Sociology and Communication Studies, and Engineering and Computer Science?*


<br>

**Articles:** 

Collins, R., & Steffen-Fluhr, N. (2019). Hidden patterns: Using social network analysis to track career trajectories of women STEM faculty. *Equality, Diversity and Inclusion: An International Journal.* https://doi.org/10.1108/EDI-09-2017-0183

Ghiasi, G., Larivière, V., & Sugimoto, C. R. (2015). On the compliance of women engineers with a gendered scientific system. *PloS one, 10*(12), e0145931. https://doi.org/10.1371/journal.pone.0145931

Jadidi, M., Karimi, F., Lietz, H., & Wagner, C. (2018). Gender disparities in science? Dropout, productivity, collaborations and success of male and female computer scientists. *Advances in Complex Systems, 21*(3-4), 1750011. https://doi.org/10.1142/S0219525917500114

*More to add...*
