---
title: "Descriptives"
#bibliography: references.bib
author: "Daphne Janssen"
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r Sys.time()`

<!---
Last compiled on `r format(Sys.time(), '%B, %Y')`
--->

<br>

----

# Descriptives Computing science staff RU

<br>

## Descriptives variables dataset 

```{r}
# start with clean workspace 
rm(list=ls())
getwd()
```
```{r}
# loading packages
library(tidyverse)
```

```{r}
# loading data

load("C:\\Users\\Daphn\\Documenten\\pubnets\\data\\names_df2_v20220106.RData")

load("C:\\Users\\Daphn\\Documenten\\pubnets\\data\\names_df_publications_v20221006.RData")
```

```{r}
# selecting the sociology sample

#wrong gs_id (just by eye-balling)
deselect <- c(24,25, 34,35)

names_df %>% 
  filter(affiliation=="Radboud University") %>%
  filter(field=="sociology") %>% 
  filter(!(id %in% deselect)) -> soc_df

names_df_publications %>% 
  filter(gs_id %in% soc_df$gs_id) -> socpub_df
  
soc_df %>% 
  select(c("gs_id", "lastname")) %>%
  right_join(socpub_df) -> socpub_df

rm(names_df)  
rm(names_df_publications)  
rm(deselect)
```

```{r}
# Cleaning the names a bit
soc_df$lastname_pubs <- as.character(str_split(soc_df$lastname, pattern=",", n = 2, simplify = TRUE)[,1]) 
socpub_df$author <- tolower(socpub_df$author)
```

```{r}
# Saving the data in a descriptives folder
save(soc_df, file="C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\soc_df.RData")
save(socpub_df, file="C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\socpub_df.RData")
```

```{r}
# Adding value labels to the variable 'dutch'
soc_df$dutch <- factor(soc_df$dutch,
levels = c(0,1),
labels = c("non-dutch", "dutch"))
```

```{r}
# Frequencies of the variables 'dutch' and 'gender'
table(soc_df$gender)
table(soc_df$dutch)
```

```{r}
# Visualisation of the gendercomposition of the three waves
library(ggplot2)

ggplot(soc_df, aes(factor(gender), fill = factor(gender))) +
    geom_bar()
```

```{r}
# Visualisation of the ethnicitycomposition of the three waves
library(ggplot2)

ggplot(soc_df, aes(factor(dutch), fill = factor(dutch))) +
    geom_bar()
```

```{r}
# Descriptives of 'gender' and 'dutch'
summary(soc_df$gender)
summary(soc_df$dutch)
```

<br>

## Array

### Creating an array

```{r}
# start with clean workspace 
rm(list=ls())
getwd()
```

```{r}
# loading packages
library(data.table) 
library(tidyverse) 
require(stringi)
require(RSiena)
```

```{r}
# loading data
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\soc_df.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\socpub_df.RData")
```

```{r}
# Network based on publications
network2016_2017 <- matrix(NA, nrow=nrow(soc_df), ncol=nrow(soc_df))
network2018_2019 <- matrix(NA, nrow=nrow(soc_df), ncol=nrow(soc_df))
network2020_2022 <- matrix(NA, nrow=nrow(soc_df), ncol=nrow(soc_df))
```

```{r}
#select publications of the corresponding time era
pubs_sel <- socpub_df %>%
    filter(year>=2016 & year<=2017)

#fill the matrix
for (ego in 1: nrow(soc_df)) {
  name_ego <- soc_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==soc_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(soc_df)){
    name_alter <- soc_df$last_name[alter] #which alter? 
    network2016_2017[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```

```{r}
#select publications of the corresponding time era
pubs_sel <- socpub_df %>%
    filter(year>=2018 & year<=2019)

#fill the matrix
for (ego in 1: nrow(soc_df)) {
  name_ego <- soc_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==soc_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(soc_df)){
    name_alter <- soc_df$last_name[alter] #which alter? 
    network2018_2019[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```

```{r}
#select publications of the corresponding time era
pubs_sel <- socpub_df %>%
    filter(year>=2020 & year<=2022)

#fill the matrix
for (ego in 1: nrow(soc_df)) {
  name_ego <- soc_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==soc_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(soc_df)){
    name_alter <- soc_df$last_name[alter] #which alter? 
    network2020_2022[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```

```{r}
# Putting all adjacency matrices in an array 
net_soc_array <- array(data = c(network2016_2017, network2018_2019, network2020_2022), dim=c(dim(network2020_2022),3))
```

```{r}
# Saving the array
save(net_soc_array, file="C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\soc_net_array.RData")
```

<br>

### Descriptives based on array

```{r}
# Start with a clean workspace 
rm(list=ls())
getwd()
```

```{r}
# Loading packages
library(data.table) 
library(tidyverse) 
require(stringi)
require(RSiena)
require(igraph)
```

```{r}
# Loading dataobjects
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\soc_net_array.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\soc_df.RData")
```

```{r}
# Degrees in a histogram
net1 <- net_soc_array[ ,, 1]
diag(net1) <- 0
degree <- rowSums(net1)
hist(degree)

net2 <- net_soc_array[ ,, 2]
diag(net2) <- 0
degree <- rowSums(net2)
hist(degree)

net3 <- net_soc_array[ ,, 3]
diag(net3) <- 0
degree <- rowSums(net3)
hist(degree)
```

```{r}
# Correcting for publications which are not present for both authors
net1 <- net_soc_array[,,1]
net1 <- net1 + t(net1)
net1[net1==2] <- 1
diag(net1) <- 0
degree <- rowSums(net1)
hist(degree)

net2 <- net_soc_array[,,2]
net2 <- net2 + t(net2)
net2[net2==2] <- 1
diag(net2) <- 0
degree <- rowSums(net2)
hist(degree)

net3 <- net_soc_array[,,3]
net3 <- net3 + t(net3)
net3[net3==2] <- 1
diag(net3) <- 0
degree <- rowSums(net3)
hist(degree)
```



<br>

## Plot

```{r}
# Igraph object 
net1_g <- igraph::graph_from_adjacency_matrix(net1, mode = c("undirected"))
plot(net1_g)

igraph::degree(net1_g)
igraph::dyad.census(net1_g)
igraph::triad.census(net1_g)
igraph::edge_density(net1_g)

sum(igraph::degree(net1_g))

net2_g <- igraph::graph_from_adjacency_matrix(net2, mode = c("undirected"))
plot(net2_g)

igraph::degree(net2_g)
igraph::dyad.census(net2_g)
igraph::triad.census(net2_g)
igraph::edge_density(net2_g)

sum(igraph::degree(net2_g))

net3_g <- igraph::graph_from_adjacency_matrix(net3, mode = c("undirected"))
plot(net3_g)

igraph::degree(net3_g)
igraph::dyad.census(net3_g)
igraph::triad.census(net3_g)
igraph::edge_density(net3_g)

sum(igraph::degree(net3_g))
```


<br>

### Changing the plot to make it more insightful

```{r}
# Making the size of the nodes proportinal to the betweenness score
V(net1_g)$size = betweenness(net1_g, normalized = T, directed = FALSE) * 60 + 10  
plot(net1_g, mode = "undirected")
V(net2_g)$size = betweenness(net2_g, normalized = T, directed = FALSE) * 60 + 10  
plot(net2_g, mode = "undirected")
V(net3_g)$size = betweenness(net3_g, normalized = T, directed = FALSE) * 60 + 10  
plot(net3_g, mode = "undirected")
```

```{r}
# Putting the information of socnet1 into a dataframe
in_network <- data.frame(as_ids(V(net1_g)))
in_network <- data.frame(as_ids(V(net2_g)))
in_network <- data.frame(as_ids(V(net3_g)))

# Giving the first column the name 'number'
names(in_network)[1] <- "number"

# Creating the object 'ru_nodes' with the names and total cites of the sociology staffmembers
ru_nodes <- soc_df[, c("name", "total_cites")]

# Adding a new variable to the dataset ru_nodes, called 'number' representing the numbers of the nodes
ru_nodes$number <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32)

# Combining in_network with ru_nodes, based on the numbers of the nodes
in_network <- left_join(in_network, ru_nodes, by = c("number" = "number"))
in_network$vcol <- ifelse(is.na(in_network$total_cites), "#E69F00", "#56B4E9")

# Makig a plot with the color of the nodes defined by citation and the names of the sociology staff members
plot(net1_g,
     vertex.color = in_network$vcol,
     vertex.label = in_network$name)

plot(net2_g,
     vertex.color = in_network$vcol,
     vertex.label = in_network$name)

plot(net3_g,
     vertex.color = in_network$vcol,
     vertex.label = in_network$name)
```

```{r}
# Defining the colors of the nodes by gender
soc_df$gcol <- ifelse(soc_df$gender == "male", "cyan4", "coral1")
plot(net1_g,
     vertex.color = soc_df$gcol,
     vertex.label = soc_df$name)

plot(net2_g,
     vertex.color = soc_df$gcol,
     vertex.label = soc_df$name)

plot(net3_g,
     vertex.color = soc_df$gcol,
     vertex.label = soc_df$name)
```

```{r}
# Making a few changes to the plot to make it more easy to derive information from it
plot(net1_g,
     vertex.color = soc_df$gcol,
     vertex.label = soc_df$name,
     vertex.size = 10,  
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 0.5) 

plot(net2_g,
     vertex.color = soc_df$gcol,
     vertex.label = soc_df$name,
     vertex.size = 10,  
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 0.5) 

plot(net3_g,
     vertex.color = soc_df$gcol,
     vertex.label = soc_df$name,
     vertex.size = 10,  
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 0.5) 
```

```{r}
# Adapting the node size to the number of citations
in_network$total_cites <- as.numeric(in_network$total_cites)

plot(net1_g,
     vertex.color = soc_df$gcol,
     vertex.label = soc_df$name,
     vertex.size = ifelse(!is.na(in_network$total_cites), log(in_network$total_cites), 2),
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 0.5) 

plot(net2_g,
     vertex.color = soc_df$gcol,
     vertex.label = soc_df$name,
     vertex.size = ifelse(!is.na(in_network$total_cites), log(in_network$total_cites), 2),
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 0.5) 

plot(net3_g,
     vertex.color = soc_df$gcol,
     vertex.label = soc_df$name,
     vertex.size = ifelse(!is.na(in_network$total_cites), log(in_network$total_cites), 2),
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 0.5) 
```

<br>

### Other descriptives based on plot

```{r}
# number of nodes: all networks consist of 36 nodes, or individuals.
vcount(net1_g)
vcount(net2_g)
vcount(net3_g)
```

```{r}
# number of edges: the number of edges increases for each wave.
ecount(net1_g)
ecount(net2_g)
ecount(net3_g)
```

```{r}
# transitivity: the number of nodes with a (high) transitivity index increases for each wave
transitivity(net1_g, type = c("localundirected"), isolates = c("NaN", "zero"))
transitivity(net2_g, type = c("localundirected"), isolates = c("NaN", "zero"))
transitivity(net3_g, type = c("localundirected"), isolates = c("NaN", "zero"))
```

```{r}
# betweennessL the amount of nodes that act as bridges increases for each wave 
igraph::betweenness(net1_g, directed = FALSE)
igraph::betweenness(net2_g, directed = FALSE)
igraph::betweenness(net3_g, directed = FALSE)
```

```{r}
# global transitivity: the global transitivity increases from wave 1 to wave 2, and decreases from wave 2 to wave 3. 
igraph::transitivity(net1_g, type = "global")
sna::gtrans(net1)

triad_net1 <- data.frame(sna::triad.census(net1))

transitivity_net1 <- (3 * triad_net1$X300)/(triad_net1$X201 + 3 * triad_net1$X300)
transitivity_net1

igraph::transitivity(net2_g, type = "global")
sna::gtrans(net2)

triad_net2 <- data.frame(sna::triad.census(net2))

transitivity_net2 <- (3 * triad_net2$X300)/(triad_net2$X201 + 3 * triad_net2$X300)
transitivity_net2

igraph::transitivity(net3_g, type = "global")
sna::gtrans(net3)

triad_net3 <- data.frame(sna::triad.census(net3))

transitivity_net3 <- (3 * triad_net3$X300)/(triad_net3$X201 + 3 * triad_net3$X300)
transitivity_net3
```

```{r}
# Defining functions

# density: observed relations divided by possible relations
fdensity <- function(x) {
    # x is your nomination network make sure diagonal cells are NA
    diag(x) <- NA
    # take care of RSiena structural zeros, set as missing.
    x[x == 10] <- NA
    sum(x == 1, na.rm = T)/(sum(x == 1 | x == 0, na.rm = T))
}

# calculate intragroup density
fdensityintra <- function(x, A) {
    # A is matrix indicating whether nodes in dyad have same node attributes
    diag(x) <- NA
    x[x == 10] <- NA
    diag(A) <- NA
    sum(x == 1 & A == 1, na.rm = T)/(sum((x == 1 | x == 0) & A == 1, na.rm = T))
}

# calculate intragroup density
fdensityinter <- function(x, A) {
    # A is matrix indicating whether nodes in dyad have same node attributes
    diag(x) <- NA
    x[x == 10] <- NA
    diag(A) <- NA
    sum(x == 1 & A != 1, na.rm = T)/(sum((x == 1 | x == 0) & A != 1, na.rm = T))
}

# construct dyadcharacteristic whether nodes are similar/homogenous
fhomomat <- function(x) {
    # x is a vector of node-covariate
    xmat <- matrix(x, nrow = length(x), ncol = length(x))
    xmatt <- t(xmat)
    xhomo <- xmat == xmatt
    return(xhomo)
}
```

```{r}
fdensity(net1)
fdensity(net2)
fdensity(net3)

net_soc_array <- array(data = c(net1, net2, net3), dim=c(dim(net1),3))
```


<br>

## Gender segregation

```{r}
# Creating a matrix for the same gender 
gender_m <-  fhomomat(soc_df$gender)
```

```{r}
# Calculating intragroup density
fdensityintra(x=net1, A=gender_m) 
fdensityintra(x=net2, A=gender_m) 
fdensityintra(x=net3, A=gender_m) 
```

```{r}
# Only selecting women
selectie <- soc_df$gender=="female"
```

```{r}
# Calculating the density for men and women seperately (intergroup density)
fdensity(net1[selectie,selectie])
fdensity(net2[selectie,selectie])
fdensity(net3[selectie,selectie])

fdensity(net1[!selectie,!selectie])
fdensity(net2[!selectie,!selectie])
fdensity(net3[!selectie,!selectie])
```


<br>



# Descriptives Computing science staff RU

<br>

## Descriptives variables dataset 

```{r}
# start with clean workspace 
rm(list=ls())
getwd()
```

```{r}
# loading packages
library(tidyverse)
```

```{r}
# loading data

load("C:\\Users\\Daphn\\Documenten\\pubnets\\data\\names_df2_v20220106.RData")

load("C:\\Users\\Daphn\\Documenten\\pubnets\\data\\names_df_publications_v20221006.RData")
```

```{r}
# selecting the computer science sample
names_df %>% 
  filter(affiliation=="Radboud University") %>%
  filter(field=="computer science") -> comp_df

names_df_publications %>% 
  filter(gs_id %in% comp_df$gs_id) -> comppub_df
  
comp_df %>% 
  select(c("gs_id", "lastname")) %>%
  right_join(comppub_df) -> comppub_df

rm(names_df)  
rm(names_df_publications)  
```

```{r}
# Cleaning the names a bit
comp_df$lastname_pubs <- as.character(str_split(comp_df$lastname, pattern=",", n = 2, simplify = TRUE)[,1]) 
comppub_df$author <- tolower(comppub_df$author)
```

```{r}
# Saving the data in a descriptives folder
save(comp_df, file="C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\comp_df.RData")
save(comppub_df, file="C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\comppub_df.RData")
```

```{r}
# Adding value labels to the variable 'dutch'
comp_df$dutch <- factor(comp_df$dutch,
levels = c(0,1),
labels = c("non-dutch", "dutch"))
```

```{r}
# Frequencies of the variables 'dutch' and 'gender'
table(comp_df$gender)
table(comp_df$dutch)
```

```{r}
# Visualisation of the gendercomposition of the three waves
library(ggplot2)

ggplot(comp_df, aes(factor(gender), fill = factor(gender))) +
    geom_bar()
```

```{r}
# Visualisation of the ethnicitycomposition of the three waves
library(ggplot2)

ggplot(comp_df, aes(factor(dutch), fill = factor(dutch))) +
    geom_bar()
```

```{r}
# Descriptives of 'gender' and 'dutch'
summary(comp_df$gender)
summary(comp_df$dutch)
```

<br>

## Array

### Creating an array

```{r}
# start with clean workspace 
rm(list=ls())
getwd()
```

```{r}
# loading packages
library(data.table) 
library(tidyverse) 
require(stringi)
require(RSiena)
```

```{r}
# loading data
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\comp_df.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\comppub_df.RData")
```

```{r}
# Network based on publications
network2016_2017 <- matrix(NA, nrow=nrow(comp_df), ncol=nrow(comp_df))
network2018_2019 <- matrix(NA, nrow=nrow(comp_df), ncol=nrow(comp_df))
network2020_2022 <- matrix(NA, nrow=nrow(comp_df), ncol=nrow(comp_df))
```

```{r}
#select publications of the corresponding time era
pubs_sel <- comppub_df %>%
    filter(year>=2016 & year<=2017)

#fill the matrix
for (ego in 1: nrow(comp_df)) {
  name_ego <- comp_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==comp_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(comp_df)){
    name_alter <- comp_df$last_name[alter] #which alter? 
    network2016_2017[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```

```{r}
#select publications of the corresponding time era
pubs_sel <- comppub_df %>%
    filter(year>=2018 & year<=2019)

#fill the matrix
for (ego in 1: nrow(comp_df)) {
  name_ego <- comp_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==comp_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(comp_df)){
    name_alter <- comp_df$last_name[alter] #which alter? 
    network2018_2019[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```

```{r}
#select publications of the corresponding time era
pubs_sel <- comppub_df %>%
    filter(year>=2020 & year<=2022)

#fill the matrix
for (ego in 1: nrow(comp_df)) {
  name_ego <- comp_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==comp_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(comp_df)){
    name_alter <- comp_df$last_name[alter] #which alter? 
    network2020_2022[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```

```{r}
# Putting all adjacency matrices in an array 
net_comp_array <- array(data = c(network2016_2017, network2018_2019, network2020_2022), dim=c(dim(network2020_2022),3))
```

```{r}
# Saving the array
save(net_comp_array, file="C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\comp_net_array.RData")
```

<br>

### Descriptives based on array

```{r}
# Start with a clean workspace 
rm(list=ls())
getwd()
```

```{r}
# Loading packages
library(data.table) 
library(tidyverse) 
require(stringi)
require(RSiena)
require(igraph)
```

```{r}
# Loading dataobjects
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\comp_net_array.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\comp_df.RData")
```

```{r}
# Degrees in a histogram
compnet1 <- net_comp_array[ ,, 1]
diag(compnet1) <- 0
degree <- rowSums(compnet1)
hist(degree)

compnet2 <- net_comp_array[ ,, 2]
diag(compnet2) <- 0
degree <- rowSums(compnet2)
hist(degree)

compnet3 <- net_comp_array[ ,, 3]
diag(compnet3) <- 0
degree <- rowSums(compnet3)
hist(degree)
```

```{r}
# Correcting for publications which are not present for both authors
compnet1 <- net_comp_array[,,1]
compnet1 <- compnet1 + t(compnet1)
compnet1[compnet1==2] <- 1
diag(compnet1) <- 0
degree <- rowSums(compnet1)
hist(degree)

compnet2 <- net_comp_array[,,2]
compnet2 <- compnet2 + t(compnet2)
compnet2[compnet2==2] <- 1
diag(compnet2) <- 0
degree <- rowSums(compnet2)
hist(degree)

compnet3 <- net_comp_array[,,3]
compnet3 <- compnet3 + t(compnet3)
compnet3[compnet3==2] <- 1
diag(compnet3) <- 0
degree <- rowSums(compnet3)
hist(degree)
```



<br>

## Plot

```{r}
# Igraph object 
compnet1_g <- igraph::graph_from_adjacency_matrix(compnet1, mode = c("undirected"))
plot(compnet1_g)

igraph::degree(compnet1_g)
igraph::dyad.census(compnet1_g)
igraph::triad.census(compnet1_g)
igraph::edge_density(compnet1_g)

sum(igraph::degree(compnet1_g))

compnet2_g <- igraph::graph_from_adjacency_matrix(compnet2, mode = c("undirected"))
plot(compnet2_g)

igraph::degree(compnet2_g)
igraph::dyad.census(compnet2_g)
igraph::triad.census(compnet2_g)
igraph::edge_density(compnet2_g)

sum(igraph::degree(compnet2_g))

compnet3_g <- igraph::graph_from_adjacency_matrix(compnet3, mode = c("undirected"))
plot(compnet3_g)

igraph::degree(compnet3_g)
igraph::dyad.census(compnet3_g)
igraph::triad.census(compnet3_g)
igraph::edge_density(compnet3_g)

sum(igraph::degree(compnet3_g))
```


<br>

### Changing the plot to make it more insightful

```{r}
# Making the size of the nodes proportinal to the betweenness score
V(compnet1_g)$size = betweenness(compnet1_g, normalized = T, directed = FALSE) * 60 + 10  
plot(compnet1_g, mode = "undirected")
V(compnet2_g)$size = betweenness(compnet2_g, normalized = T, directed = FALSE) * 60 + 10  
plot(compnet2_g, mode = "undirected")
V(compnet3_g)$size = betweenness(compnet3_g, normalized = T, directed = FALSE) * 60 + 10  
plot(compnet3_g, mode = "undirected")
```

```{r}
# Putting the information of soccompnet1 into a dataframe
in_network <- data.frame(as_ids(V(compnet1_g)))
in_network <- data.frame(as_ids(V(compnet2_g)))
in_network <- data.frame(as_ids(V(compnet3_g)))

# Giving the first column the name 'number'
names(in_network)[1] <- "number"

# Creating the object 'ru_nodes' with the names and total cites of the sociology staffmembers
ru_nodes <- comp_df[, c("name", "total_cites")]

# Adding a new variable to the dataset ru_nodes, called 'number' representing the numbers of the nodes
ru_nodes$number <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61,	62,	63,	64,	65,	66,	67,	68,	69, 70,	71,	72,	73, 74,	75,	76,	77,	78,	79, 80,	81,	82,	83,	84,	85,	86,	87,	88,	89, 90,	91,	92,	93,	94,	95,	96,	97,	98,	99, 100, 101,	102, 103,	104,	105,	106,	107,	108,	109, 110,	111,	112,	113,	114,	115)

# Combining in_network with ru_nodes, based on the numbers of the nodes
in_network <- left_join(in_network, ru_nodes, by = c("number" = "number"))
in_network$vcol <- ifelse(is.na(in_network$total_cites), "#E69F00", "#56B4E9")

# Makig a plot with the color of the nodes defined by citation and the names of the sociology staff members
plot(compnet1_g,
     vertex.color = in_network$vcol,
     vertex.label = in_network$name)

plot(compnet2_g,
     vertex.color = in_network$vcol,
     vertex.label = in_network$name)

plot(compnet3_g,
     vertex.color = in_network$vcol,
     vertex.label = in_network$name)
```

```{r}
# Defining the colors of the nodes by gender
comp_df$gcol <- ifelse(comp_df$gender == "male", "cyan4", "coral1")
plot(compnet1_g,
     vertex.color = comp_df$gcol,
     vertex.label = comp_df$name)

plot(compnet2_g,
     vertex.color = comp_df$gcol,
     vertex.label = comp_df$name)

plot(compnet3_g,
     vertex.color = comp_df$gcol,
     vertex.label = comp_df$name)
```

```{r}
# Making a few changes to the plot to make it more easy to derive information from it

plot(compnet1_g,
     vertex.color = comp_df$gcol,
     vertex.label = comp_df$name,
     vertex.size= 10,
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 25) 

plot(compnet2_g,
     vertex.color = comp_df$gcol,
     vertex.label = comp_df$name,
     vertex.size = 10,  
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 0.5) 

plot(compnet3_g,
     vertex.color = comp_df$gcol,
     vertex.label = comp_df$name,
     vertex.size = 10,  
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 0.5) 
```

```{r}
# Adapting the node size to the number of citations
in_network$total_cites <- as.numeric(in_network$total_cites)

plot(compnet1_g,
     vertex.color = comp_df$gcol,
     vertex.label = comp_df$name,
     vertex.size = ifelse(!is.na(in_network$total_cites), log(in_network$total_cites), 2),
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 0.5) 

plot(compnet2_g,
     vertex.color = comp_df$gcol,
     vertex.label = comp_df$name,
     vertex.size = ifelse(!is.na(in_network$total_cites), log(in_network$total_cites), 2),
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 0.5) 

plot(compnet3_g,
     vertex.color = comp_df$gcol,
     vertex.label = comp_df$name,
     vertex.size = ifelse(!is.na(in_network$total_cites), log(in_network$total_cites), 2),
     vertex.frame.color = "gray",  
     vertex.label.color = "black",  
     vertex.label.family = "Helvetica", 
     vertex.label.cex = 0.7,  
     vertex.label.dist = 0.8,  
     edge.curved = 0.2,  
     edge.arrow.size = 0.5) 
```

<br>

### Other descriptives based on plot

```{r}
# number of nodes: all networks consist of 115 nodes, or individuals.
vcount(compnet1_g)
vcount(compnet2_g)
vcount(compnet3_g)
```

```{r}
# number of edges: the number of edges increases for each wave.
ecount(compnet1_g)
ecount(compnet2_g)
ecount(compnet3_g)
```

```{r}
# transitivity: the number of nodes with a (high) transitivity index increases for each wave
transitivity(compnet1_g, type = c("localundirected"), isolates = c("NaN", "zero"))
transitivity(compnet2_g, type = c("localundirected"), isolates = c("NaN", "zero"))
transitivity(compnet3_g, type = c("localundirected"), isolates = c("NaN", "zero"))
```

```{r}
# betweennessL the amount of nodes that act as bridges increases for each wave 
igraph::betweenness(compnet1_g, directed = FALSE)
igraph::betweenness(compnet2_g, directed = FALSE)
igraph::betweenness(compnet3_g, directed = FALSE)
```

```{r}
# global transitivity: the global transitivity increases from wave 1 to wave 2, and decreases from wave 2 to wave 3. 
igraph::transitivity(compnet1_g, type = "global")
sna::gtrans(compnet1)

triad_compnet1 <- data.frame(sna::triad.census(compnet1))

transitivity_compnet1 <- (3 * triad_compnet1$X300)/(triad_compnet1$X201 + 3 * triad_compnet1$X300)
transitivity_compnet1

igraph::transitivity(compnet2_g, type = "global")
sna::gtrans(compnet2)

triad_compnet2 <- data.frame(sna::triad.census(compnet2))

transitivity_compnet2 <- (3 * triad_compnet2$X300)/(triad_compnet2$X201 + 3 * triad_compnet2$X300)
transitivity_compnet2

igraph::transitivity(compnet3_g, type = "global")
sna::gtrans(compnet3)

triad_compnet3 <- data.frame(sna::triad.census(compnet3))

transitivity_compnet3 <- (3 * triad_compnet3$X300)/(triad_compnet3$X201 + 3 * triad_compnet3$X300)
transitivity_compnet3
```

```{r}
# Defining functions

# density: observed relations divided by possible relations
fdensity <- function(x) {
    # x is your nomination network make sure diagonal cells are NA
    diag(x) <- NA
    # take care of RSiena structural zeros, set as missing.
    x[x == 10] <- NA
    sum(x == 1, na.rm = T)/(sum(x == 1 | x == 0, na.rm = T))
}

# calculate intragroup density
fdensityintra <- function(x, A) {
    # A is matrix indicating whether nodes in dyad have same node attributes
    diag(x) <- NA
    x[x == 10] <- NA
    diag(A) <- NA
    sum(x == 1 & A == 1, na.rm = T)/(sum((x == 1 | x == 0) & A == 1, na.rm = T))
}

# calculate intragroup density
fdensityinter <- function(x, A) {
    # A is matrix indicating whether nodes in dyad have same node attributes
    diag(x) <- NA
    x[x == 10] <- NA
    diag(A) <- NA
    sum(x == 1 & A != 1, na.rm = T)/(sum((x == 1 | x == 0) & A != 1, na.rm = T))
}

# construct dyadcharacteristic whether nodes are similar/homogenous
fhomomat <- function(x) {
    # x is a vector of node-covariate
    xmat <- matrix(x, nrow = length(x), ncol = length(x))
    xmatt <- t(xmat)
    xhomo <- xmat == xmatt
    return(xhomo)
}
```

```{r}
fdensity(compnet1)
fdensity(compnet2)
fdensity(compnet3)

net_comp_array <- array(data = c(compnet1, compnet2, compnet3), dim=c(dim(compnet1),3))
```


<br>

## Gender segregation

```{r}
# Creating a matrix for the same gender 
gender_m <-  fhomomat(comp_df$gender)
```

```{r}
# Calculating intragroup density
fdensityintra(x=compnet1, A=gender_m) 
fdensityintra(x=compnet2, A=gender_m) 
fdensityintra(x=compnet3, A=gender_m) 
```

```{r}
# Only selecting women
selectie <- comp_df$gender=="female"
```

```{r}
# Calculating the density for men and women seperately (intergroup density)
fdensity(compnet1[selectie,selectie])
fdensity(compnet2[selectie,selectie])
fdensity(compnet3[selectie,selectie])

fdensity(compnet1[!selectie,!selectie])
fdensity(compnet2[!selectie,!selectie])
fdensity(compnet3[!selectie,!selectie])
```


