---
title: "Introduction and Research Question"
#bibliography: references.bib
author: "Daphne Janssen"
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r Sys.time()`

<!---
Last compiled on `r format(Sys.time(), '%B, %Y')`
--->

<br>

----

# Introduction and Research question

For a long time, up to the present day, women have been underrepresented in STEM (science, technology, engineering and mathematics) studies and professions. According to the website *stemwomen.com*, the gap between men and women taking STEM courses is widest in computer sciences and engingeering & technology: 19 percent of students are women, compared to 81 percent who are men. Moreover, research shows that men account for 80 percent of the total amount of scientific production within engineering. Although women publish their papers in higher impact journals, they get less citated by the scientific engineering community compared to men (Ghiasi, Larivière & Sugimoto,2015).

Previous research shows that co-authorship is highly relevant to promotion and productivity (Collins & Steffen-Fluhr, 2019). However, it is known that there are gender differences in scientific collaborations, putting women at a disadvantage (Jadidi, Karimi, Lietz, & Wagner,2018). While some studies find that because of the homophily principle, men in STEM studies are more likely to collaborate with other men than with women, thus reducing the chances for women to be signalled (Collins & Steffen-Fluhr, 2019), other studies conclude that women practice more gender homophily (Jadidi et al., 2018). Ghiasi et al. (2015) find that engineers predominantly collaborate with men, contributing to scientific structures recognised by male dominance. Jadidi et al. (2018) further find that women are generally less likely to enter into collaborations that will lead to success and are more likely to be embedded in ego networks with structural holes. 


I would like to contribute to the current scientific field concerning gender differences in social networks within STEM, by investigating to what extent men and women differ in who they collaborate with and on which topics they publish their articles, and to what extent this differs between sociology and engineering within three Dutch universities.

<br> 

Based on the above I derived the following **research question**: 
*To what extent is there a difference between men and women within Sociology and Engineering in (1) who they work with and (2) which topics they study compared to three universities within the Netherlands?*

<br>

**Articles:** 

Collins, R., & Steffen-Fluhr, N. (2019). Hidden patterns: Using social network analysis to track career trajectories of women STEM faculty. *Equality, Diversity and Inclusion: An International Journal.* https://doi.org/10.1108/EDI-09-2017-0183

Ghiasi, G., Larivière, V., & Sugimoto, C. R. (2015). On the compliance of women engineers with a gendered scientific system. *PloS one, 10*(12), e0145931. https://doi.org/10.1371/journal.pone.0145931

Jadidi, M., Karimi, F., Lietz, H., & Wagner, C. (2018). Gender disparities in science? Dropout, productivity, collaborations and success of male and female computer scientists. *Advances in Complex Systems, 21*(3-4), 1750011. https://doi.org/10.1142/S0219525917500114


