---
title: "RSiena"
#bibliography: references.bib
author: "Daphne Janssen"
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r Sys.time()`

<!---
Last compiled on `r format(Sys.time(), '%B, %Y')`
--->

<br>

----

# RSiena analysis Sociology staff RU


## Preparations
```{r}
# Starting with a clean workspace 
rm(list=ls())
```

```{r}
# Requiring packages 
library(data.table) 
library(tidyverse) 
require(stringi)
require(RSiena)
```

```{r}
# Loading dataobjects
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\soc_net_array.RData")
#load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_collabs.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Data\\soc_staff_cit.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\soc_df.RData")
```

```{r}
# Selecting the dependent variable
net <- sienaDependent(net_soc_array)
?sienaDependent
```

<br>

## Step 1: define data
```{r}
# Gender
gender <- as.numeric(soc_df$gender=="female")
gender <- coCovar(gender)
```

```{r}
# Ethnicity
dutch <- as.numeric(soc_df$dutch==1)
dutch <- coCovar(dutch)
```

```{r}
# Number of collaborators
#ncollabs <- soc_df$ncollabs
#ncollabs <- coCovar(ncollabs)

#class(ncollabs)
```

```{r}
# Publications per wave for each staff computing science staff member
pubsw1 <- pubsw2 <- pubsw3 <- NA

for (i in 1:length(soc_df$gs_id)) {
  pubsw1[i] <- nrow(soc_staff_cit[(soc_staff_cit$gs_id == soc_df$gs_id[i]) & soc_staff_cit$year>=2016 & soc_staff_cit$year<=2017,])
  pubsw2[i] <- nrow(soc_staff_cit[(soc_staff_cit$gs_id == soc_df$gs_id[i]) & soc_staff_cit$year>=2018 & soc_staff_cit$year<=2019,])
  pubsw3[i] <- nrow(soc_staff_cit[(soc_staff_cit$gs_id == soc_df$gs_id[i]) & soc_staff_cit$year>=2020 & soc_staff_cit$year<=2022,])
}

pub_df <- as.matrix(data.frame(pubsw1, pubsw2, pubsw3))

pubs <- varCovar(pub_df)
```

```{r}
# Year first publication
soc_staff_cit %>% group_by(gs_id) %>%
  mutate(pub_first = min(year)) %>%
  select(c("gs_id", "pub_first")) %>%
  distinct(gs_id, pub_first, .keep_all = TRUE) -> firstpub_df

soc_df <- left_join(soc_df, firstpub_df)

#if no publication yet, set pub_first op 2023
soc_df %>% mutate(pub_first = replace_na(pub_first, 2023)) -> soc_df

pub_first <-  coCovar(soc_df$pub_first)

mydata <- sienaDataCreate(net, gender, dutch, pubs, pub_first)
```

<br>

## Step 2: create effects structure
```{r}
myeff <- getEffects(mydata)
effectsDocumentation(myeff)
```

<br>

## Step 3: get initial description
```{r}
print01Report(mydata, modelname = "C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\soc_init")
```

<br>

## Step 4: specify model
```{r}
myeff <- includeEffects(myeff, density)
myeff <- includeEffects(myeff, isolateNet)
myeff <- includeEffects(myeff, transTrip)
myeff <- includeEffects(myeff, transTies)
myeff <- includeEffects(myeff, egoPlusAltX)
myeff <- includeEffects(myeff, sameX, interaction1 = "gender")
myeff <- includeEffects(myeff, sameX, interaction1 = "dutch")
myeff <- includeEffects(myeff, egoX, interaction1 = "pub_first")
myeff <- includeEffects(myeff, egoX, interaction1 = "gender")
myeff <- includeEffects(myeff, egoX, interaction1 = "dutch")
```

```{r}
# Possible structural effects:
# density
# transTrip, transTies, gwesp, 
# inPop, outAct, inAct of outPop, 
# isolateNet
# degPlus of egoPlusAltX
```

<br>

## Step 5: estimate
```{r}
myAlgorithm <- sienaAlgorithmCreate(projname = "soc_init")
```

```{r, eval=FALSE}
(ans6 <- siena07(myAlgorithm, data = mydata, effects = myeff))
# (the outer parentheses lead to printing the obtained result on the screen) if necessary, estimate
# further
# (ans <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns = ans))

save(ans6, file="C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\soc_model1")
```

```{r, echo=FALSE}
#load("C:\\Users\\Daphn\\Documenten\\labjournal\\soc_init.txt")
  
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\soc_model1")
```

![](results/soc_init.txt){#id .class width=100% height=200px}


<br>


# RSiena analysis Computing science staff RU

<br>

## Preparations
```{r}
# Starting with a clean workspace 
rm(list=ls())
```

```{r}
# Requiring packages 
library(data.table) 
library(tidyverse) 
require(stringi)
require(RSiena)
```

```{r}
# Loading dataobjects
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\comp_net_array.RData")
load("C:\\Users\\Daphn\\Documenten\\pubnets\\Data\\names_staff_cit_v20221006.RData")
load("C:\\Users\\Daphn\\Documenten\\labjournal\\data\\descriptives\\comp_df.RData")
```

```{r}
# Selecting the dependent variable
net <- sienaDependent(net_comp_array)
?sienaDependent
```

<br>

## Step 1: define data
```{r}
# Gender
gender <- as.numeric(comp_df$gender=="female")
gender <- coCovar(gender)
```

```{r}
# Ethnicity
dutch <- as.numeric(comp_df$dutch==1)
dutch <- coCovar(dutch)
```

```{r}
# Number of collaborators
#ncollabs <- comp_df$ncollabs
#ncollabs <- coCovar(ncollabs)

#class(ncollabs)
```

```{r}
# Creating a dataset with information on the google scholar id's, publications and citations of each computing science staff member 

names_staff_cit %>%
  filter(names_staff_cit$gs_id %in% comp_df$gs_id) -> comp_staff_cit

# Removing duplicates
sel <- (duplicated.data.frame(comp_staff_cit))

comp_staff_cit %>% 
  filter(!sel) -> comp_staff_cit
```

```{r}
# Publications per wave for each staff computing science staff member
pubsw1 <- pubsw2 <- pubsw3 <- NA

for (i in 1:length(comp_df$gs_id)) {
  pubsw1[i] <- nrow(comp_staff_cit[(comp_staff_cit$gs_id == comp_df$gs_id[i]) & comp_staff_cit$year>=2016 & comp_staff_cit$year<=2017,])
  pubsw2[i] <- nrow(comp_staff_cit[(comp_staff_cit$gs_id == comp_df$gs_id[i]) & comp_staff_cit$year>=2018 & comp_staff_cit$year<=2019,])
  pubsw3[i] <- nrow(comp_staff_cit[(comp_staff_cit$gs_id == comp_df$gs_id[i]) & comp_staff_cit$year>=2020 & comp_staff_cit$year<=2022,])
}

pub_df <- as.matrix(data.frame(pubsw1, pubsw2, pubsw3))

pubs <- varCovar(pub_df)
```

```{r}
# Year of first publication
comp_staff_cit %>% group_by(gs_id) %>%
  mutate(pub_first = min(year)) %>%
  select(c("gs_id", "pub_first")) %>%
  distinct(gs_id, pub_first, .keep_all = TRUE) -> firstpub_df

comp_df <- left_join(comp_df, firstpub_df)

#if no publication yet, set pub_first op 2023
comp_df %>% mutate(pub_first = replace_na(pub_first, 2023)) -> comp_df

pub_first <-  coCovar(comp_df$pub_first)

mydata <- sienaDataCreate(net, gender, dutch, pubs, pub_first)
```

<br>

## Step 2: create effects structure
```{r}

myeff <- getEffects(mydata)
effectsDocumentation(myeff)
```

<br>

## Step 3: get initial description
```{r}
print01Report(mydata, modelname = "C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_init")
```

<br>

## Step 4: specify model
```{r}
myeff <- includeEffects(myeff, density)
myeff <- includeEffects(myeff, isolateNet)
myeff <- includeEffects(myeff, transTrip)
myeff <- includeEffects(myeff, transTies)
myeff <- includeEffects(myeff, egoPlusAltX)
myeff <- includeEffects(myeff, sameX, interaction1 = "gender")
myeff <- includeEffects(myeff, sameX, interaction1 = "dutch")
myeff <- includeEffects(myeff, egoX, interaction1 = "pub_first")
myeff <- includeEffects(myeff, egoX, interaction1 = "gender")
myeff <- includeEffects(myeff, egoX, interaction1 = "dutch")
```

```{r}
# Possible structural effects:
# density
# transTrip, transTies, gwesp, 
# inPop, outAct, inAct of outPop, 
# isolateNet
# degPlus of egoPlusAltX
```

<br>

## Step 5: estimate
```{r}
myAlgorithm <- sienaAlgorithmCreate(projname = "comp_init")
```

```{r, eval=FALSE}
(ans1 <- siena07(myAlgorithm, data = mydata, effects = myeff))
# (the outer parentheses lead to printing the obtained result on the screen) if necessary, estimate
# further
# (ans <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns = ans))

save(ans1, file="C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_model1")
```

```{r, echo=FALSE}
#load("C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_init")
  
load("C:\\Users\\Daphn\\Documenten\\labjournal\\Results\\comp_model1")
```

![](results/comp_init.txt){#id .class width=100% height=200px}