---
title: "Research Question 6"
#bibliography: references.bib
author: "Daphne Janssen"
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r Sys.time()`

<!---
Last compiled on `r format(Sys.time(), '%B, %Y')`
--->

<br>

----


# Research Question(s) week 6

<br>

*1) To what extent does the application of the gender homophily principle to the publication network of male and female scientists at Radboud University within Sociology and Computer Science change between 2016 and 2022?*

*2) To what extent is there a difference in the application of the gender homophily principle to the publication network of male and female scientists at Radboud University, between Sociology and Computing science?*

*3) To what extent do structural network effects influence the difference in gender homophily in the publication network of male and female scientists at Radboud University, within Sociology and Computing science?*

