---
title: "Research Question 4"
#bibliography: references.bib
author: "Daphne Janssen"
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r Sys.time()`

<!---
Last compiled on `r format(Sys.time(), '%B, %Y')`
--->

<br>

----


# Research Question week 4

<br>

*1)	To what extent does the gender homophily principle apply to the publication network of male and female scientists at Radboud University, within Sociology and Computing science?*

*2)	To what extent is there a difference in the application of the gender homophily principle to the publication network of male and female scientists at Radboud University, between Sociology and Computing science?*

*3)	To what extent do structural network effects influence the difference in gender homophily in the publication network of male and female scientists at Radboud University, within Sociology and Computing science?*
