
```{r}
#start with clean workspace 
rm(list=ls())
getwd()
```

# packages

```{r}
library(data.table) 
library(tidyverse) 
require(stringi)
require(Rsiena)

```

# load data

```{r}
load("./data/names_df2_v20220106.RData")
load("./data/names_df_publications_v20221006.RData")
```

# sample selection sociology

```{r}

#wrong gs_id (just by eye-balling)
#deselect <- c(24,25, 34,35)

names_df %>% 
  filter(affiliation=="Radboud University") %>%
  filter(field=="sociology") %>% 
  filter(!(id %in% deselect)) -> soc_df

names_df_publications %>% 
  filter(gs_id %in% soc_df$gs_id) -> socpub_df
  
soc_df %>% 
  select(c("gs_id", "lastname")) %>%
  right_join(socpub_df) -> socpub_df

rm(names_df)  
rm(names_df_publications)  
rm(deselect)
```
<!--- to do, deze selecties opslaan ---> 


# clean names a bit

```{r}
soc_df$lastname_pubs <- as.character(str_split(soc_df$lastname, pattern=",", n = 2, simplify = TRUE)[,1]) 

socpub_df$author <- tolower(socpub_df$author)

```

```{r}
save(soc_df, file="./data/descriptives/soc_df.RData")
save(socpub_df, file="./data/descriptives/socpub_df.RData")
```

# network based on publications

NATURALLY, THIS IS JUST AN EXAMPLE. fOR rSIENA, YOU NEED AT LEAST 3 NETWORKS. tHUS YOU HAVE TO TWEAK THE PERIOD AS YOU DEEM FIT. 

```{r}
network2016_2017 <- matrix(NA, nrow=nrow(soc_df), ncol=nrow(soc_df))
network2018_2019 <- matrix(NA, nrow=nrow(soc_df), ncol=nrow(soc_df))
network2020_2022 <- matrix(NA, nrow=nrow(soc_df), ncol=nrow(soc_df))
```

```{r}
#select publications of the corresponding time era
pubs_sel <- socpub_df %>%
    filter(year>=2016 & year<=2017)

# to do: we gebruiken nu str_detect, maar wrs moeten we een exact match gebruiken. 

#fill the matrix
for (ego in 1: nrow(soc_df)) {
  name_ego <- soc_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==soc_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(soc_df)){
    name_alter <- soc_df$last_name[alter] #which alter? 
    network2016_2017[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```

```{r}
#select publications of the corresponding time era
pubs_sel <- socpub_df %>%
    filter(year>=2018 & year<=2019)

# to do: we gebruiken nu str_detect, maar wrs moeten we een exact match gebruiken. 

#fill the matrix
for (ego in 1: nrow(soc_df)) {
  name_ego <- soc_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==soc_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(soc_df)){
    name_alter <- soc_df$last_name[alter] #which alter? 
    network2018_2019[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```

```{r}
#select publications of the corresponding time era
pubs_sel <- socpub_df %>%
    filter(year>=2020 & year<=2022)

# to do: we gebruiken nu str_detect, maar wrs moeten we een exact match gebruiken. 

#fill the matrix
for (ego in 1: nrow(soc_df)) {
  name_ego <- soc_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==soc_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(soc_df)){
    name_alter <- soc_df$last_name[alter] #which alter? 
    network2020_2022[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```


# netwerken undirected maken, dat zijn ze grappig genoeg niet! 

alle adjacency matrix in een array stoppen. 
```{r}
net_soc_array <- array(data = c(network2016_2017, network2018_2019, network2020_2022), dim=c(dim(network2020_2022),3))

```


```{r}
save(net_soc_array, file="./data/descriptives/soc_net_array.RData")
```



# computer science

```{r}
#start with clean workspace 
rm(list=ls())
getwd()
```

# packages

```{r}
library(data.table) 
library(tidyverse) 
require(stringi)
require(Rsiena)

```

# load data

```{r}
load("./data/names_df2_v20220106.RData")
load("./data/names_df_publications_v20221006.RData")
```


# sample selection computer science

```{r}

#wrong gs_id (just by eye-balling)
#deselect <- c(24,25, 34,35)
#filter(!(id %in% deselect))

names_df %>% 
  filter(affiliation=="Radboud University") %>%
  filter(field=="computer science")  -> soc_df

names_df_publications %>% 
  filter(gs_id %in% soc_df$gs_id) -> socpub_df
  
soc_df %>% 
  select(c("gs_id", "lastname")) %>%
  right_join(socpub_df) -> socpub_df

rm(names_df)  
rm(names_df_publications)  
rm(deselect)
```
<!--- to do, deze selecties opslaan ---> 


# clean names a bit

```{r}
soc_df$lastname_pubs <- as.character(str_split(soc_df$lastname, pattern=",", n = 2, simplify = TRUE)[,1]) 

socpub_df$author <- tolower(socpub_df$author)

```

```{r}
cs_df <- soc_df
cspub_df <- socpub_df

save(cs_df, file="./data/descriptives/cs_df.RData")
save(cspub_df, file="./data/descriptives/cspub_df.RData")
```

# network based on publications

NATURALLY, THIS IS JUST AN EXAMPLE. fOR rSIENA, YOU NEED AT LEAST 3 NETWORKS. tHUS YOU HAVE TO TWEAK THE PERIOD AS YOU DEEM FIT. 

```{r}
network2016_2017 <- matrix(NA, nrow=nrow(cs_df), ncol=nrow(cs_df))
network2018_2019 <- matrix(NA, nrow=nrow(cs_df), ncol=nrow(cs_df))
network2020_2022 <- matrix(NA, nrow=nrow(cs_df), ncol=nrow(cs_df))
```

```{r}
#select publications of the corresponding time era
pubs_sel <- cspub_df %>%
    filter(year>=2016 & year<=2017)

# to do: we gebruiken nu str_detect, maar wrs moeten we een exact match gebruiken. 

#fill the matrix
for (ego in 1: nrow(cs_df)) {
  name_ego <- cs_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==cs_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(cs_df)){
    name_alter <- cs_df$last_name[alter] #which alter? 
    network2016_2017[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```

```{r}
#select publications of the corresponding time era
pubs_sel <- cspub_df %>%
    filter(year>=2018 & year<=2019)

# to do: we gebruiken nu str_detect, maar wrs moeten we een exact match gebruiken. 

#fill the matrix
for (ego in 1: nrow(cs_df)) {
  name_ego <- cs_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==cs_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(cs_df)){
    name_alter <- cs_df$last_name[alter] #which alter? 
    network2018_2019[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```

```{r}
#select publications of the corresponding time era
pubs_sel <- cspub_df %>%
    filter(year>=2020 & year<=2022)

# to do: we gebruiken nu str_detect, maar wrs moeten we een exact match gebruiken. 

#fill the matrix
for (ego in 1: nrow(soc_df)) {
  name_ego <- soc_df$lastname_pubs[ego] #which ego? 
  pubs_sel2 <- pubs_sel[pubs_sel$gs_id==soc_df$gs_id[ego],] #selecteer de publicaties van ego
  for (alter in 1:nrow(soc_df)){
    name_alter <- soc_df$last_name[alter] #which alter? 
    network2020_2022[ego,alter] <- as.numeric(sum(str_detect(pubs_sel2$author, name_alter)) > 1)  #did alter publish with ego
  }
}
```


# netwerken undirected maken, dat zijn ze grappig genoeg niet! 

alle adjacency matrix in een array stoppen. 
```{r}
net_cs_array <- array(data = c(network2016_2017, network2018_2019, network2020_2022), dim=c(dim(network2020_2022),3))

```


```{r}
save(net_cs_array, file="./data/descriptives/cs_net_array.RData")
```



