
# sociology 

```{r}
#start with clean workspace 
rm(list=ls())
getwd()
```

# packages

```{r}
library(data.table) 
library(tidyverse) 
require(stringi)
require(RSiena)

```

# load data

```{r}
#load dataobjects
load("./data/descriptives/soc_net_array.RData")
load("./data/descriptives/soc_df.RData")

```

zorg er voor dat de netwerken symmetrisch zijn

```{r}
net1 <- net_soc_array[,,1]
net1 <- net1 + t(net1)
net1[net1==2] <- 1

net2 <- net_soc_array[,,2]
net2 <- net2 + t(net2)
net2[net2==2] <- 1

net3 <- net_soc_array[,,3]
net3 <- net3 + t(net3)
net3[net3==2] <- 1


net_soc_array <- array(data = c(net1, net2, net3), dim=c(dim(net1),3))

```


```{r}
#dependent
net <- sienaDependent(net_soc_array)
```


```{r}
### Step 1: define data
#gender
gender <- as.numeric(soc_df$gender=="female")
gender <- coCovar(gender)
```


```{r, eval=FALSE}
#ncollabs
ncollabs <- (soc_df$ncollabs)
ncollabs <- coCovar(ncollabs)

#pubs: time-varying
pubsw1 <- pubsw2 <- pubsw3 <- pubsw4 <- NA
for (i in 1:length(soc_df$gs_id)) {
  pubsw1[i] <- nrow(soc_staff_cit[(soc_staff_cit$gs_id == soc_df$gs_id[i]) & soc_staff_cit$year>=2010 & soc_staff_cit$year<=2012,])
  pubsw2[i] <- nrow(soc_staff_cit[(soc_staff_cit$gs_id == soc_df$gs_id[i]) & soc_staff_cit$year>=2013 & soc_staff_cit$year<=2015,])
  pubsw3[i] <- nrow(soc_staff_cit[(soc_staff_cit$gs_id == soc_df$gs_id[i]) & soc_staff_cit$year>=2016 & soc_staff_cit$year<=2018,])
  pubsw4[i] <- nrow(soc_staff_cit[(soc_staff_cit$gs_id == soc_df$gs_id[i]) & soc_staff_cit$year>=2019 & soc_staff_cit$year<=2021,])
}

pub_df <- as.matrix(data.frame(pubsw1, pubsw2, pubsw3, pubsw4))
pubs <- varCovar(pub_df)

#year first pub
soc_staff_cit %>% group_by(gs_id) %>%
  mutate(pub_first = min(year)) %>% 
  select(c("gs_id", "pub_first")) %>%
  distinct(gs_id, pub_first, .keep_all = TRUE) -> firstpub_df

soc_df <- left_join(soc_df, firstpub_df)

#if no publication yet, set pub_first op 2023
soc_df %>% mutate(pub_first = replace_na(pub_first, 2023)) -> soc_df

pub_first <-  coCovar(soc_df$pub_first)
```


```{r}
mydata <- sienaDataCreate(net, gender)
```


```{r}
### Step 2: create effects structure
myeff <- getEffects(mydata)
#effectsDocumentation(myeff)
```


```{r}
### Step 3: get initial description
print01Report(mydata, modelname = "./results/soc_init")
```

![](results/soc_init.txt){#id .class width=100% height=200px}

optie 1

```{r}
knitr::incl
```


```{r, eval=FALSE}
### Step4: specify model
myeff <- includeEffects(myeff, degPlus) #some publish a lot, some not. (interpretation: talent/luck? )
myeff <- includeEffects(myeff, transTriads)
myeff <- includeEffects(myeff, sameX, interaction1 = "gender")
#myeff <- includeEffects(myeff, egoPlusAltX, interaction1 = "pubs") #if you have many pubs, you will probably publish more per year? 
myeff <- includeEffects(myeff, absDiffX, interaction1 = "pub_first") #senior publishes with junior

### Step5 estimate
myAlgorithm <- sienaAlgorithmCreate(projname = "soc_init")
(ans <- siena07(myAlgorithm, data = mydata, effects = myeff))
# (the outer parentheses lead to printing the obtained result on the screen) if necessary, estimate
# further
(ans <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns = ans))
```


# computer science (RU)

```{r}
#start with clean workspace 
rm(list=ls())
getwd()
```

# packages

```{r}
library(data.table) 
library(tidyverse) 
require(stringi)
require(RSiena)

```

# load data

```{r}
#load dataobjects
load("./data/descriptives/cs_net_array.RData")
load("./data/descriptives/cs_df.RData")

```

zorg er voor dat de netwerken symmetrisch zijn

```{r}
net1 <- net_cs_array[,,1]
net1 <- net1 + t(net1)
net1[net1==2] <- 1

net2 <- net_cs_array[,,2]
net2 <- net2 + t(net2)
net2[net2==2] <- 1

net3 <- net_cs_array[,,3]
net3 <- net3 + t(net3)
net3[net3==2] <- 1


net_cs_array <- array(data = c(net1, net2, net3), dim=c(dim(net1),3))

```


```{r}
#dependent
net <- sienaDependent(net_cs_array)
```

gender
```{r}
### Step 1: define data
#gender
gender <- as.numeric(cs_df$gender=="female")
gender <- coCovar(gender)
```

cites per jaar

Let op. de cites zijn al geaggregeerd per jaar. Als je dus publicaties per jaar zou willen tellen, moet je ander bestand gebruiken!! 

```{r}
load("./data/names_staff_cit_v20221006.RData")

names_staff_cit %>%
  filter(names_staff_cit$gs_id %in% cs_df$gs_id) -> cs_staff_cit

#duplicates verwijderen
sel <- (duplicated.data.frame(cs_staff_cit))

cs_staff_cit %>% 
  filter(!sel) -> cs_staff_cit


```


```{r, eval=FALSE}
#cites: time-varying
citesw1 <- citesw2 <- citesw3 <- NA
for (i in 1:length(cs_df$gs_id)) {
  citesw1[i] <- sum(cs_staff_cit[(cs_staff_cit$gs_id == cs_df$gs_id[i]) & cs_staff_cit$year>=2016 & cs_staff_cit$year<=2017,c("cites")])
  
  citesw2[i] <- sum(cs_staff_cit[(cs_staff_cit$gs_id == cs_df$gs_id[i]) & cs_staff_cit$year>=2018 & cs_staff_cit$year<=2019,c("cites")])
  
  citesw3[i] <- sum(cs_staff_cit[(cs_staff_cit$gs_id == cs_df$gs_id[i]) & cs_staff_cit$year>=2020 & cs_staff_cit$year<=2022,c("cites")])
}

#given the right-skewed distribution, perhaps better to use log-transformation. 
#to avoid value -Inf I added .5 to everyone
cites_df <- log(as.matrix(data.frame(citesw1, citesw2, citesw3)) + .5)
cites <- varCovar(cites_df)
```

year first pub

```{r, eval=FALSE}
#year first pub
cs_staff_cit %>% group_by(gs_id) %>%
  mutate(pub_first = min(year)) %>% 
  select(c("gs_id", "pub_first")) %>%
  distinct(gs_id, pub_first, .keep_all = TRUE) -> firstpub_df

cs_df <- left_join(cs_df, firstpub_df)

#if no publication yet, set pub_first op 2023
cs_df %>% mutate(pub_first = replace_na(pub_first, 2023)) -> cs_df

pub_first <-  coCovar(cs_df$pub_first)
```


```{r}
mydata <- sienaDataCreate(net, gender, cites, pub_first)
```


```{r}
### Step 2: create effects structure
myeff <- getEffects(mydata)
#effectsDocumentation(myeff)
```


```{r}
### Step 3: get initial description
print01Report(mydata, modelname = "./results/soc_init")
```

<!---to show this file in your lab journal use the following code---> 

![](results/soc_init.txt){#id .class width=100% height=200px}



```{r, eval=FALSE}
### Step4: specify model
myeff <- includeEffects(myeff, degPlus) #some publish a lot, some not. (interpretation: talent/luck? )
myeff <- includeEffects(myeff, transTriads)
myeff <- includeEffects(myeff, sameX, interaction1 = "gender")
myeff <- includeEffects(myeff, egoPlusAltX, interaction1 = "cites") #if you have many cites, you will get more collaboration within department 
myeff <- includeEffects(myeff, egoPlusAltX, interaction1 = "pub_first") #if you are older, you will get more collaboration within department 
```


```{r, eval=FALSE}
### Step5 estimate
myAlgorithm <- sienaAlgorithmCreate(projname = "cs_init")
(ans <- siena07(myAlgorithm, data = mydata, effects = myeff))
# (the outer parentheses lead to printing the obtained result on the screen) if necessary, estimate
# further
(ans <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns = ans))
```


